### 2015-04-20 analysis/e5_analysis.Rmd
e5 analysis

PSYCH 130 - EXPERIMENT 4 ANALYSIS
=================================

Preliminaries.

```{r}
rm(list=ls())
source("analysis/useful_dplyr.R")
library("rjson")
```

Read in files and consolidate to the same directory.

```{r}
files <- dir("production-results/e5/", pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf <- paste("production-results/e5/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- data.frame(workerid = jd$WorkerId, 
                   scale = jd$answer$scale,
                   degree = jd$answer$degree,
                   stars = jd$answer$manipulation_level,
                   judgment = as.numeric(jd$answer$judgment),
                   language = jd$answer$language)
  d.raw <- bind_rows(d.raw, id)
}
```

Holding off on exclusions - everyone put English, and training 2 was confusing.

```{r}
# compliant <- d.raw %>% 
#   filter(grepl("training", scale)) %>%
#   group_by(workerid) %>%
#   summarise(compliant = judgment[scale == "training1" & degree == "hi"] > 3 & 
#               judgment[scale == "training1" & degree == "low"] < 3)
#   
# d <- d.raw %>% 
#   left_join(compliant) %>%
#   filter(grepl(pattern = "english", 
#                language, ignore.case = TRUE), 
#          domain != "training1", 
#          domain != "training2", 
#          compliant == TRUE)
d <- d.raw %>% filter(scale != "training1")
```

Main Analysis
-------------

First look at means broken down by epistemic condition.

```{r}
ms <- d %>%
  mutate(stars = as.numeric(as.character(stars))/20) %>%
  group_by(scale, degree, stars) %>%
  summarise(cil = ci.low(judgment), 
            cih = ci.high(judgment), 
            judgment = mean(judgment))

qplot(x = stars, y = judgment, col = degree,
      geom = "line", stat = "identity", 
      position = "dodge", 
      data = ms) + 
  geom_linerange(aes(ymin = judgment - cil, 
                     ymax = judgment + cih)) + 
  facet_wrap(~scale)
```