### 2015-05-15 models/Model_Playspace.Rmd
model revision

---
title: "Scalar_M1"
author: "BPeloquin"
date: "April 23, 2015"
output: html_document
---

Preliminaries.

```{r}
rm(list=ls())
source("analysis/useful_dplyr.R")
library(rjson)
```

## Notes
* <none, some, all> condition: the model correctly produces implicature.
* However "all"" condition is predicted to be compabitible with a 4 rating.
* This appears to be driven by several factors
  * 1) High priors over 4
  * 2) Fairly high compabitibiliy with All and four stars from e5

## Basic form of the model

$$ p_{L_p} (r \mid u) \propto \frac{p_S(u | r) p_{L_0}(r)}{\sum_{u' \in alts}{p_S(u' \mid r) p_{L_0}(r)}}$$

Read Data 
---------

```{r}
priors = read.csv("models/prior.csv")
speaker = read.csv("models/speaker.csv")
listener = read.csv("models/listener.csv")
```

## MODEL IMPLEMENTATION

Key:
* u = utterance [string]
* alts = alternative utterances
* r = rating index [1-5]
* d.info = informativity judgements

```{r}
#cost function in length of chars
# u.cost = function(u) {
#   return(nchar(u))
# }

#speaker likelihood helper 
speaker.lhd = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost)) #Rethink alpha and costs?
}

#speaker likelihood normalized
speaker.prob = function(rating, degree, m, alpha) {
  num = speaker.lhd(m[rating, degree], alpha)
  norm = 0
  for (i in 1:ncol(m)) {
    norm = norm + speaker.lhd(m[rating, i], alpha)
  }
  return(num / norm)
}

#non-normalized posterior
nn.post = function(rating, degree, m, alpha, useprior) {

  if (useprior) {
    prior <- priors[rating, "prior.p"]
  } else {
    prior = .2 
  }

  speaker.prob(rating, degree, m, alpha) * prior
}

#normalized posterior
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  norm = 0
  for (i in 1:5) {
    norm = norm + nn.post(i, degree, m, alpha, useprior)
  }
  return(nn / norm)
}  
```

model wrapper functions:

```{r}
run.rsa <- function(d, alpha = 1, useprior = TRUE, usenone = FALSE) {

  mat <- d %>%
    select(stars, degree, speaker.p) %>%
    spread(degree, speaker.p) %>%
    mutate(hi = hi / sum(hi), 
           low = low / sum(low)) %>%
    select(hi, low)

  if (usenone) {
    mat$none <- c(1, 0, 0, 0, 0)
  }

  d$pred <- round(as.numeric(mapply(norm.post, 
                                    d$stars, 
                                    d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
} 
```

RUN MODEL
--------

Run model on actual data

```{r}
data <- left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

run it.

```{r}
md <- data %>%
  do(run.rsa(., alpha = 1, useprior = FALSE, usenone = TRUE))

qplot(stars, listener.p, col=degree, 
      data = md) + 
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 2)
```

## DATA / PLOTS

Priors
```{r}
qplot(c(1:5), priors$prior.p,
      geom="bar", stat = "identity",
      binwidth = 1, main="Priors over ratings")
```

SOME
```{r}
d.some.unif.priors = sapply(1:5, FUN=function(i){
  norm.post("some", i, scale.matrix, unif.priors)})
d.some.raw.priors = sapply(1:5, FUN=function(i){
  norm.post("some", i, scale.matrix, raw.priors)})

qplot(c(1:5), d.some.raw.priors,
      geom="bar", stat = "identity",
      binwidth = 1, main="Some Posteriors")
```

ALL
```{r}
d.all.raw.priors = sapply(1:5, FUN=function(i){
  norm.post("all", i, scale.matrix, raw.priors)})
d.all.unif.priors = sapply(1:5, FUN=function(i){
  norm.post("all", i, n.scale.matrix, unif.priors)})

qplot(c(1:5), d.all.raw.priors,
      geom="bar", stat = "identity",
      binwidth = 1, main="All Posteriors, Raw priors")
```