### 2015-07-24 analysis/e5_analysis2.Rmd
Literal listener semantics anaysis #2: Looking at matching between likert ratings and stars shown, visualized a couple ways. Looks like there is a good deal of matching, especially at extremes. Rolling together [1,2], 3, [4,5] seeing similar results as with means. Can we approximate gaussians with mean and variance for agreements here?

---
output: pdf_document
---
PSYCH 130 - EXPERIMENT 4 ANALYSIS
=================================

Preliminaries.

```{r}

rm(list=ls())
setwd("~/Desktop/Projects/scalar_implicature/")
source("analysis/useful_dplyr.R")
library(tidyr)
library(dplyr)
library("rjson")
```

Read in files and consolidate to the same directory.

```{r}
files <- dir("production-results/e5/", pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf <- paste("production-results/e5/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- data.frame(workerid = jd$WorkerId, 
                   scale = jd$answer$scale,
                   degree = jd$answer$degree,
                   stars = jd$answer$manipulation_level,
                   judgment = as.numeric(jd$answer$judgment),
                   language = jd$answer$language)
  d.raw <- bind_rows(d.raw, id)
}

# Number of participants
length(unique(d.raw$workerid))
```

Holding off on exclusions - everyone put English, and training 2 was confusing. 
```{r}
# compliant <- d.raw %>% 
#   filter(grepl("training", scale)) %>%
#   group_by(workerid) %>%
#   summarise(compliant = judgment[scale == "training1" & degree == "hi"] > 3 & 
#               judgment[scale == "training1" & degree == "low"] < 3)
#   
# d <- d.raw %>% 
#   left_join(compliant) %>%
#   filter(grepl(pattern = "english", 
#                language, ignore.case = TRUE), 
#          domain != "training1", 
#          domain != "training2", 
#          compliant == TRUE)
d <- d.raw %>% filter(scale != "training1")
```

Main Analysis
-------------
```{r}
names(d)
dim(d)
# removing language - main data.frame()
d$stars = as.numeric(as.character(d$stars))/20
dt = d[, -6]
scalars = c("some_all", "liked_loved", "palatable_delicious",
            "good_excellent", "memorable_unforgettable")
dgrs = c("low", "hi")
```

Histogram Plots for our scalars
```{r}
scalar.agreements = function(scalar_pair, degree) {
  pair = dt[which(dt$scale == scalar_pair), ]
  scalar = pair[which(pair$degree == degree), ]
  scalar = spread(scalar, stars, judgment)
  # plots
  par(mfrow = c(2,3))
  wrd = unlist(strsplit(scalar_pair, "_"))
  wrd = ifelse(degree == "low", wrd[1], wrd[2])
  likert.plots(scalar, wrd)
}
likert.plots = function(scalar, wrd) {
  barplot( table(scalar$"1"), main = paste(wrd, ": *"),
           xlab="Agree", ylab="Counts")
  barplot( table(scalar$"2"), main = paste(wrd, ": **"),
           xlab="Agree", ylab="Counts")
  barplot( table(scalar$"3"), main = paste(wrd, ": ***"),
           xlab="Agree", ylab="Counts")
  barplot( table(scalar$"4"), main = paste(wrd, ": ****"),
           xlab="Agree", ylab="Counts")
  barplot( table(scalar$"5"), main = paste(wrd, ": *****"),
           xlab="Agree", ylab="Counts")
}
for (i in scalars) {
  for (j in dgrs) {
    scalar.agreements(i, j)
  }
}
```

Matching in data
```{r}
match = dt$stars == dt$judgment
par(mfrow=c(1,1))
barplot(c("T" = mean(match), "F" = (1 - mean(match))), main = "Percent matches (judgements and stars")
dm = data.frame(dt, match)

# Matches by stars
table(select(dm, stars, match))

# Matches by stars & degrees
table(select(dm, stars, match, degree))
table(select(dm, stars, match, degree))
```

Rolling 1,2 4,5
```{r}
ds = d[, -6] %>%
  group_by(scale, degree, stars) %>%
  summarise(speaker.hi = sum( ifelse( (judgment > 3), 1, 0)),
            speaker.mid = sum( ifelse( (judgment == 3), 1, 0)),
            speaker.lo = sum( ifelse( (judgment < 3), 1, 0)))

compatibility = function(scalar_pair, degree) {
  dr = ds[which(ds$scale == scalar_pair & ds$degree == degree),] %>%
    select(stars, speaker.hi)
  return(dr$speaker.hi / sum(dr$speaker.hi))
}
plot.pairs = function(scalar_pair) {
  plot(compatibility(scalar_pair, "hi"), type="b", col="red",
       main=toString(scalar_pair), ylab="% Agree Norm", xlab="Stars",
       xlim = c(1,5), ylim = c(0,1))
  lines(compatibility(scalar_pair, "low"), type="b", col="blue")  
}
compatibility("good_excellent", "hi")
par(mfrow=c(1,1))
plot.pairs("good_excellent")
par(mfrow=c(2,3))
for (i in 1:length(scalars)) {
  plot.pairs(scalars[i])  
}
```