### 2015-09-17 analysis/e8_analysis.Rmd
general update after exp 8 run

---
title: "e8_analysis"
author: "BPeloquin"
date: "August 6, 2015"
output: html_document
---

Prelims
-------
```{r}
rm(list=ls())
setwd("~/Desktop/Projects/scalar_implicature/")
source("analysis/useful_dplyr.R")
library("rjson")
```

Read in data
------------
```{r}
files <- dir("production-results/e8/", pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf <- paste("production-results/e8/", f, sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- data.frame(workerid = jd$WorkerId,
                   scale = jd$answers$data$scale,
                   degree = jd$answers$data$degree,
                   stars = jd$answers$data$manipulation_level,
                   judgment = as.numeric(jd$answers$data$judgment),
                   language = jd$answers$data$language)
  d.raw <- bind_rows(d.raw, id)
}
```

Initial data clean
------------------
```{r}
d.raw = d.raw[c(-1, -2), ]       # remove first two rows
d = d.raw %>% filter(scale != "training1")

ms <- d %>%
  mutate(stars = as.numeric(as.character(stars))/20) %>%
  group_by(scale, degree, stars) %>%
  summarise(cil = ci.low(judgment), 
            cih = ci.high(judgment), 
            cnt.judgment = sum(judgment),
            speaker.p = mean(judgment)) 
# write to csv for model
write.csv(ms, "models/e8.csv", row.names=FALSE)
```

Literal Listener semantics plot - averages
```{r}
pdf("plots/e8_L0.pdf", width=8, height = 5)
qplot(x = stars, y = speaker.p, col = degree,
      geom = "line", stat = "identity", 
      position = "dodge", main = "Average", data = ms) + 
  geom_linerange(aes(ymin = speaker.p - cil, 
                     ymax = speaker.p + cih)) + 
  facet_wrap(~scale)
dev.off()
```

Literal Listener semantics plot - counts
```{r}
qplot(x = stars, y = cnt.judgment,
      geom = "bar", fill = factor(degree), stat = "identity", 
      position = "dodge", main = "counts", data = ms) +
  facet_wrap(~scale)
```

Looking at mean and var
```{r}

weighted.avg = function(scalar, degree) {
  return( sum(
    ms[which(ms$scale == scalar & ms$degree == degree), ]$speaker.p *
      ms[which(ms$scale == scalar & ms$degree == degree), ]$stars) / 5)   
}
weighted.avg("some_all", "low")

normals = matrix(ncol = 4, nrow = 5)
colnames(overall.cnts) = dgrs
for (i in 1:length(scalars)) {
  normals[i, 1] = 
    mean(ms[which(ms$scale == scalars[i] & ms$degree == "low"),]$cnt.judgment)
  normals[i, 2] = 
    mean(ms[which(ms$scale == scalars[i] & ms$degree == "hi"),]$cnt.judgment)
  normals[i, 3] = 
    var(ms[which(ms$scale == scalars[i] & ms$degree == "low"),]$cnt.judgment)
  normals[i, 4] = 
    var(ms[which(ms$scale == scalars[i] & ms$degree == "hi"),]$cnt.judgment)
}
colnames(normals) = c("mean hi", )
```

Looking at counts
```{r}
scalars = c("some_all",
            "liked_loved",
            "palatable_delicious",
            "good_excellent",
            "memorable_unforgettable")
dgrs = c("low", "hi")
overall.cnts = matrix(ncol = 2, nrow = 5)
colnames(overall.cnts) = dgrs
for (i in 1:length(scalars)) {
  overall.cnts[i, 1] = 
    sum(ms[which(ms$scale == scalars[i] & ms$degree == "low"),]$cnt.judgment)
  overall.cnts[i, 2] = 
    sum(ms[which(ms$scale == scalars[i] & ms$degree == "hi"),]$cnt.judgment)
}
overall.cnts = data.frame(overall.cnts)
overall.cnts$scalar = scalars

# for plotting
dc = overall.cnts %>%
  gather(degree, cnt, -scalar)
dc

qplot(scalars, y=cnt,
      geom="bar", stat="identity",
      data=dc) +
  facet_wrap(~degree)
qplot(scalars, y=hi,
      geom="bar", stat="identity", main="low", ylim=c(0, 150),
      data=data.frame(overall.cnts))

```