### 2015-09-21 analysis/e8_analysis.Rmd
added norming section - do we take care of this in RSA??? Next step to plot the literal semantics

---
title: "e8_analysis"
author: "BPeloquin"
date: "August 6, 2015"
output: html_document
---

Prelims
-------
```{r}
rm(list=ls())
setwd("~/Desktop/Projects/scalar_implicature/")
source("analysis/useful_dplyr.R")
library("rjson")
```

Read in data
------------
```{r}
files <- dir("production-results/e8/", pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf <- paste("production-results/e8/", f, sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- data.frame(workerid = jd$WorkerId,
                   scale = jd$answers$data$scale,
                   degree = jd$answers$data$degree,
                   stars = jd$answers$data$manipulation_level,
                   judgment = as.numeric(jd$answers$data$judgment),
                   language = jd$answers$data$language)
  d.raw <- bind_rows(d.raw, id)
}
```

Initial data clean
------------------
```{r}
d.raw = d.raw[c(-1, -2), ]       # remove first two rows
d = d.raw %>% filter(scale != "training1")

ms <- d %>%
  mutate(stars = as.numeric(as.character(stars))/20) %>%
  group_by(scale, degree, stars) %>%
  summarise(cil = ci.low(judgment), 
            cih = ci.high(judgment), 
            cnt.judgment = sum(judgment),
            speaker.p = mean(judgment)) 
# write to csv for model
write.csv(ms, "models/e8.csv", row.names=FALSE)
```

Literal Listener semantics plot - averages
```{r}
pdf("plots/e8_L0.pdf", width=8, height = 5)
qplot(x = stars, y = speaker.p, col = degree,
      geom = "line", stat = "identity", 
      position = "dodge", main = "Average", data = ms) + 
  geom_linerange(aes(ymin = speaker.p - cil, 
                     ymax = speaker.p + cih)) + 
  facet_wrap(~scale)
dev.off()
```

Literal Listener semantics plot - counts
```{r}
qplot(x = stars, y = cnt.judgment,
      geom = "bar", fill = factor(degree), stat = "identity", 
      position = "dodge", main = "counts", data = ms) +
  facet_wrap(~scale)
```

Looking at mean and var
```{r}

weighted.avg = function(scalar, degree) {
  return( sum(
    ms[which(ms$scale == scalar & ms$degree == degree), ]$speaker.p *
      ms[which(ms$scale == scalar & ms$degree == degree), ]$stars) / 5)   
}
weighted.avg("some_all", "low")

normals = matrix(ncol = 4, nrow = 5)
colnames(overall.cnts) = dgrs
for (i in 1:length(scalars)) {
  normals[i, 1] = 
    mean(ms[which(ms$scale == scalars[i] & ms$degree == "low"),]$cnt.judgment)
  normals[i, 2] = 
    mean(ms[which(ms$scale == scalars[i] & ms$degree == "hi"),]$cnt.judgment)
  normals[i, 3] = 
    var(ms[which(ms$scale == scalars[i] & ms$degree == "low"),]$cnt.judgment)
  normals[i, 4] = 
    var(ms[which(ms$scale == scalars[i] & ms$degree == "hi"),]$cnt.judgment)
}
colnames(normals) = c("mean hi", )
```

Looking at counts
```{r}
scalars = c("some_all",
            "liked_loved",
            "palatable_delicious",
            "good_excellent",
            "memorable_unforgettable")
dgrs = c("low", "hi")
overall.cnts = matrix(ncol = 2, nrow = 5)
colnames(overall.cnts) = dgrs
for (i in 1:length(scalars)) {
  overall.cnts[i, 1] = 
    sum(ms[which(ms$scale == scalars[i] & ms$degree == "low"),]$cnt.judgment)
  overall.cnts[i, 2] = 
    sum(ms[which(ms$scale == scalars[i] & ms$degree == "hi"),]$cnt.judgment)
}
overall.cnts = data.frame(overall.cnts)
overall.cnts$scalar = scalars

# for plotting
dc = overall.cnts %>%
  gather(degree, cnt, -scalar)
dc

qplot(scalars, y=cnt,
      geom="bar", stat="identity",
      data=dc) +
  facet_wrap(~degree)
qplot(scalars, y=hi,
      geom="bar", stat="identity", main="low", ylim=c(0, 150),
      data=data.frame(overall.cnts))
```

```{r}
# working with vector
normed.counts = NA
s = scalars[1]
d = dgrs[2]

for (s in scalars) {
  cnts = data.frame()
  for (d in dgrs) {
    cnts$scale = rep(s, 5)
    cnts$dgrs = rep(d, 5)
    cnts = ms[ms$scale == s & ms$degree == d, "cnt.judgment"] + 1       # Laplace smooth
    cnts = cnts / sum(cnts)
  }
  new.sems = c(new.sems, cnts)
}
as.vector(unlist(new.sems))
ms

# init.df
# ------
# initialze a new nrows * ncols data.frame()
init.df = function(ncols, nrows, names = NA) {
  new.df = data.frame(matrix(ncol = ncols, nrow = nrows))
  colnames(new.df) = names
  return(new.df)
}

# working wit df()
new.sems = data.frame(scale=character(),
                      degree=character(),
                      stars=numeric(),
                      cnt.judgment=numeric())
for (s in scalars) {
  cnts = init.df(4, 5, names=c("scale", "degree", "stars", "cnt.judgment"))
  for (d in dgrs) {
    cnts$scale = rep(s, 5)
    cnts$degree = rep(d, 5)
    cnts$stars = seq(1, 5)
    cnts$cnt.judgment = ms[ms$scale == s & ms$degree == d, ]$cnt.judgment + 1 # Laplace smooth
    cnts$cnt.judgment = cnts$cnt.judgment / sum(cnts$cnt.judgment)
    new.sems = rbind(new.sems, cnts)
  }
}
# for merging with ms df()
new.sems = rbind(new.sems, data.frame(scale=rep("training1", 5),
                           degree=c(rep("hi",5), rep("low",5)),
                           cnt.judgment=rep(NA,10)))
ma = merge(ms, new.sems, by=c("scale", "degree", "stars"))
```