### 2015-10-03 analysis/e9_altsAnalysis.Rmd
preliminary analylsis of alternatives in R. Using table to derive counts. Think of a better way to do this.

---
title: "e9_altsAnalysis"
author: "BPeloquin"
date: "October 2, 2015"
output: html_document
---

```{r}
#library(rjson)
rm(list = ls())
library(jsonlite)
setwd("~/Desktop/Projects/scalar_implicature/")
```

```{r}
d0 = fromJSON(file="alts.json")
names(d0)
scalar_data = c()
for (scalar in names(d0)) {
  for(word in d[[scalar]]) {
    scalar = c(scalar, word)
    print(scalar)
  }
  scalar_data = rbind(scalar_data, scalar)
}
scalar_data
```

```{r}
d1 = fromJSON("alts.json", simplifyDataFrame=T)
d1 = as.data.frame(d1)
# conver to lower case
sapply(seq(1, ncol(d1)), FUN=function(i) {
  d1[,i] = tolower(d1[,i])
})

sort(table(d1$all))
sort(table(d1$some))
sort(table(d1$unforgettable))
sort(table(d1$memorable))
sort(table(d1$excellent))
sort(table(d1$good))
sort(table(d1$loved))
sort(table(d1$liked))
sort(table(d1$delicious))
sort(table(d1$palatable))

```