### 2015-10-06 analysis/e9_altsAnalysis.Rmd
alts analysis in R, just reading in data and superficial look at alt sets

---
title: "e9_altsAnalysis"
author: "BPeloquin"
date: "October 2, 2015"
output: html_document
---

```{r}
rm(list = ls())
#library(rjson)
library(jsonlite)
fp = "~/Desktop/Projects/scalar_implicature/analysis/"
setwd("~/Desktop/Projects/scalar_implicature/")
```

**Analysis of alternative gathering experiment**

Data cleaning in alts_analysis.py

Explore
--------
```{r}
d1 = fromJSON("analysis/edited_alts.json", simplifyDataFrame=T)
d1 = as.data.frame(d1)

length(sort(table(d1$all)))
length(sort(table(d1$some)))

length(sort(table(d1$unforgettable)))
length(sort(table(d1$memorable)))

length(sort(table(d1$excellent)))
length(sort(table(d1$good)))

length(sort(table(d1$loved)))
length(sort(table(d1$liked)))

length(sort(table(d1$delicious)))
length(sort(table(d1$palatable)))

```