### 2015-10-13 analysis/e9_altsAnalysis.Rmd
organize data into top ten alterntives by scalar pairs

---
title: "e9_altsAnalysis"
author: "BPeloquin"
date: "October 2, 2015"
output: html_document
---

```{r}
rm(list = ls())
#library(rjson)
library(jsonlite)
fp = "~/Desktop/Projects/scalar_implicature/analysis/"
setwd("~/Desktop/Projects/scalar_implicature/")
```

**Analysis of alternative gathering experiment**

Note: Data cleaning in alts_analysis.py

Explore
--------
```{r}
d1 = fromJSON("analysis/edited_alts.json", simplifyDataFrame=T)
d1 = as.data.frame(d1, stringsAsFactors = FALSE)
```

```{r}
library(tidyr)
library(dplyr)
d <- d1 %>% 
  select(-high, -low) %>%
  gather(base, alt, 
         palatable, liked, good, loved, some, all, 
         delicious, excellent, memorable, unforgettable) %>%
  mutate(base = as.character(base)) %>%
  rowwise %>%
  mutate(scale = ifelse(base == "some" | base == "all", "some_all", 
                        ifelse(base == "palatable" | base == "delicious", "palatable_delicious",
                               ifelse(base == "liked" | base == "loved", "liked_loved",
                                      ifelse(base == "good" | base == "excellent", 
                                             "good_excellent", "memorable_unforgettable"))))) %>%
  group_by(scale, alt) %>%
  summarise(n = n())

d.topTen = d %>%
  filter(n >= 10)
```

**Output to text file for reference:**
```{r}
d.topTen =  d.topTen[order(d.topTen$scale, -d.topTen$n),]
write.table(d.topTen, "analysis/top_alts_output.txt")
```