### 2015-10-17 analysis/e11_analysis.Rmd
adding copy of e10 analysis for e11

---
output: pdf_document
---

Experiment e11 - Pragmatic Listener experiment with alternatives
```{r}

rm(list=ls())
setwd("~/Desktop/Projects/scalar_implicature/")
source("analysis/useful_dplyr.R")
library(tidyr)
library(dplyr)
library("rjson")

# commentary on weird RMarkdown bug
# https://github.com/hadley/xml2/issues/34
```

Data prep
---------
```{r}

# make sure to set directory path to 'production-results/e11/'
results_path = "sandbox-results/e11/"
files <- dir(results_path, pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf <- paste(results_path,f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- data.frame(workerid = jd$WorkerId, 
                   scale = jd$answers$data$scale,
                   degree = jd$answers$data$degree,
                   stars = jd$answers$data$manipulation_level,
                   judgment = as.numeric(jd$answers$data$judgment),
                   language = jd$answers$data$language)
  d.raw <- bind_rows(d.raw, id)
}

# Number of participants
length(unique(d.raw$workerid))
```

Holding off on exclusions (but exclusion code here)
---------------------------------------------------
```{r}
d <- d.raw %>% filter(scale != "training1")
# compliant <- d.raw %>% 
#   filter(grepl("training", scale)) %>%
#   group_by(workerid) %>%
#   summarise(compliant = judgment[scale == "training1" & degree == "hi"] > 3 & 
#               judgment[scale == "training1" & degree == "low"] < 3)
#   
# d <- d.raw %>% 
#   left_join(compliant) %>%
#   filter(grepl(pattern = "english", 
#                language, ignore.case = TRUE), 
#          domain != "training1", 
#          domain != "training2", 
#          compliant == TRUE)
```

Main analysis below
-------------------

Let's remove training runs
```{r}
d.raw = d.raw[c(-1, -2), ]       # remove first two rows
d = d.raw %>% filter(scale != "training1")
```

Populate a new df
```{r}
ms <- d %>%
  mutate(stars = as.numeric(as.character(stars))/20) %>%
  group_by(scale, degree, stars) %>%
  summarise(cil = ci.low(judgment), 
            cih = ci.high(judgment), 
            cnt.judgment = sum(judgment),
            speaker.p = mean(judgment)) 
```

Plot our literal listener semantic (counts data)
```{r}
pdf("plots/e10_L0Counts.pdf", width=8, height = 5)
qplot(x = stars, y = cnt.judgment, col = degree,
      geom = "line", stat = "identity", 
      position = "dodge", 
      data = ms) + 
  geom_linerange(aes(ymin = cnt.judgment - cil, 
                     ymax = cnt.judgment + cih)) + 
  facet_wrap(~scale) +
  ggtitle("Literal listener counts")
dev.off()

pdf("plots/e10_L0Percent.pdf", width=8, height = 5)
qplot(x = stars, y = speaker.p, col = degree,
      geom = "line", stat = "identity", 
      position = "dodge", 
      data = ms) + 
  geom_linerange(aes(ymin = speaker.p - cil, 
                     ymax = speaker.p + cih)) + 
  facet_wrap(~scale) +
  ggtitle("Literal listener percent 'yes'")
dev.off()
```

```{r}
# prop.calc = function(scalar, degree, stars, df) {
#   total = sum(df[(df$scale == scalar) & (df$degree == degree), "cnt.judgment"])
#   prop = df[(df$scale == scalar) & (df$degree == degree) & (df$stars = stars), "cnt.judgment"]
#   return(prop / total)
# }
# scales = rep("some_all", 5)
# stars = seq(1:5)
# degree = rep("hi1", 5)
# cnt.judgment = c(0, 0, 0, 2, 28)
# 
# df = data.frame(scales, stars, degree, cnt.judgment)
# prop.calc("some_all", "hi1", 1, df)
# 
# ms$judgment.prop = ms$judgment
# # write to csv for model
# #write.csv(ms, "models/e10.csv", row.names=FALSE)

```