### 2015-10-20 analysis/e11_analysis.Rmd
checking for compliance of training trials - appears we had 100% compliance on e11

---
output: pdf_document
---

Experiment e11 - Pragmatic Listener experiment with alternatives
```{r}
rm(list=ls())
setwd("~/Desktop/Projects/scalar_implicature/")
source("analysis/useful_dplyr.R")
library(tidyr)
library(dplyr)
library("rjson")
library(ggplot2)
# commentary on weird RMarkdown bug
# https://github.com/hadley/xml2/issues/34
```

Data prep
---------
```{r}

# make sure to set directory path to 'production-results/e11/'
results_path = "production-results/e11/"
files <- dir(results_path, pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf <- paste(results_path,f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- data.frame(workerid = jd$WorkerId, 
                   scale = jd$answer$data$scale,
                   degree = jd$answer$data$degree,
                   stars = as.numeric(jd$answer$data$judgment),
                   language = jd$answer$data$language,
                   gender = jd$answers$data$gender,
                   age = jd$answers$data$age)
  d.raw <- bind_rows(d.raw, id)
}

# Number of participants
length(unique(d.raw$workerid))
```

Holding off on exclusions (but exclusion code here)
---------------------------------------------------
```{r}
# Make sure people answered training trials correctly
compliant <- d.raw %>% 
   filter(grepl("training", scale)) %>%
   group_by(workerid) %>%
   summarise(compliant = stars[scale == "training1" & degree == "hi2"] > 3 & 
               stars[scale == "training1" & degree == "low1"] < 3)

d <- d.raw %>% 
  left_join(compliant) %>%
  filter(grepl(pattern = "english", 
               language, ignore.case = TRUE), 
         scale != "training1", 
         scale != "training2", 
         compliant == TRUE)

# remove training runs
d <- d.raw %>% filter(scale != "training1")
head(d)
```

Main analysis below
-------------------

Populate a new df
```{r}
density <- d %>%
  group_by(degree, scale) %>%
  mutate(n = n()) %>%
  group_by(degree, scale, stars) %>%
  summarise(listener.p = n() / n[1])
```

Plot our pragmatic listener judgments
```{r}
write.csv(density, "models/e11.csv", row.names=FALSE)
```

```{r}
pdf("plots/e11.pdf", width = 8, height = 4)
qplot(stars, listener.p, facets = degree ~ scale, 
      geom="bar", stat = "identity",
      binwidth = 1, data=density)
dev.off()
```