### 2015-10-30 models/RSA+AmbiguousSemantics.Rmd
spelling

---
title: "RSA + Scalar implicature / TBD"
author: "BPeloquin/MFrank"
date: "October 29, 2015"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
    toc_depth: 2
---

# Pragmatics / implicature introduction

## Rational Speech-act theory

## Extending RSA to scalar implicature

# Experimental investigations / star rating paradigm

## Literal listener studies

### Description

### Results

## Pragmatic listener studies

### Description

### Results

# Formalizing Grice / RSA / Bayesian model

## Current model description

## Implementation

# Connecting experiments and model

## Baseline model performance

## Miniumum alternatives model

## Eliciting alternatives experiment

### Description

### Results

## Full model (alternatives) performance

#Conclusions