### 2015-11-03 analysis/exp_data_clean.Rmd
Data cleaning for pragmatic listener experiments

---
title: "data_cleaning"
author: "BPeloquin"
date: "November 1, 2015"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
---

Data cleaning code for Literal listener and Pragmatic listener experiments
```{r load_data, results="hide", message=FALSE, warning=FALSE}
rm(list = ls())
setwd("/Users/benpeloquin/Desktop/Projects/scalar_implicature/analysis")
source("useful_dplyr.R")
library(rjson)
```

# Literal Listener study e8 output

## Read in data
```{r read_data, message=FALSE, warning=FALSE}
# Grab raw data from .json
# ------------------------
results_path = "../production-results/e8/"
files = dir(results_path, pattern = "*.json")
d.raw = data.frame()

for (f in files) {
  jf = paste(results_path, f, sep="")
  jd = fromJSON(paste(readLines(jf), collapse=""))
  id = data.frame(workerid = jd$WorkerId,
                   scale = jd$answers$data$scale,
                   degree = jd$answers$data$degree,
                   stars = jd$answers$data$manipulation_level,
                   judgment = as.numeric(jd$answers$data$judgment),
                   language = jd$answers$data$language)
  d.raw = bind_rows(d.raw, id)
}

str(d.raw)
head(d.raw)
```

## Compliance checks

I'm holding off on compliance for e8. We made it a bit easier in e10 (both were agreement) and I think we're missing a large portion of the sample unncessarily.

1. training1:high  -> presented 5 stars asked if the person thought the food deserved a `high` rating

2. training:low    -> presented 4 stars, asked if the person thought the food deserved a `low` rating
```{r}
# Sample size pre compliance check
n.pre.compliance = length(unique(d.raw$workerid))
# Check for compliance
compliant <- d.raw %>% 
  filter(grepl("training1", scale)) %>%
  group_by(workerid) %>%
  summarise(compliant = judgment[scale == "training1" & degree == "hi"] == 1 & 
              judgment[scale == "training1" & degree == "low"] == 0)

# Sample size post compliance check
n.post.compliance = sum(compliant$compliant)

# If set to TRUE then enroce compliance for e8 (n reduced to 16)
if (FALSE) { # Enforce compliance check for training 1
  d <- d.raw %>% 
  left_join(compliant) %>%
  filter(grepl(pattern = "english", 
               language, ignore.case = TRUE), 
         scale != "training1", 
         compliant == TRUE)
} else { # Only enforce language check
  d <- d.raw %>% 
  left_join(compliant) %>%
  filter(grepl(pattern = "english", 
               language, ignore.case = TRUE), 
         scale != "training1")
}

# Current sample
n.post.compliance = length(unique(d$workerid)); n.post.compliance
```

## Cleaning / formatting - output
```{r}
# Preliminary data cleaning
# -------------------------
ms = d %>%
  mutate(stars = as.numeric(as.character(stars))/20) %>%
  group_by(scale, degree, stars) %>%
  summarise(cil = ci.low(judgment), 
            cih = ci.high(judgment), 
            cnt.judgment = sum(judgment),
            speaker.p = mean(judgment))

str(ms)
head(ms)
# Output to CSV
# --------------
# write.csv(ms, "../models/L0_e8.csv", row.names=FALSE)
```

# Literal Listener study e10 output

## Read in data
```{r read_data2, message=FALSE, warning=FALSE}
results_path = "../production-results/e10/"
files <- dir(results_path, pattern = "*.json")
d.raw <- data.frame()

for (f in files) {
  jf = paste(results_path,f,sep="")
  jd = fromJSON(paste(readLines(jf), collapse=""))
  id = data.frame(workerid = jd$WorkerId, 
                   scale = jd$answers$data$scale,
                   degree = jd$answers$data$degree,
                   stars = jd$answers$data$manipulation_level,
                   judgment = as.numeric(jd$answers$data$judgment),
                   language = jd$answers$data$language)
  d.raw = bind_rows(d.raw, id)
}

str(d.raw)
head(d.raw)
```

## Compliance checks

Enforce compliance in e10. This check seemed more reasonable:

1. training1:high2  -> presented 5 stars asked if the person thought the food deserved a `high(est)` rating

2. training:low1    -> presented 1 star, asked if the person thought the food deserved a `low(est)` rating
```{r}
# Sample size pre compliance check
n.pre.compliance = length(unique(d.raw$workerid)); n.pre.compliance

compliant <- d.raw %>% 
  filter(grepl("training", scale)) %>%
  group_by(workerid) %>%
  summarise(compliant = (judgment[scale == "training1" & degree == "hi2"] == 1) && 
              (judgment[scale == "training1" & degree == "low1"] == 1))

num_uncompliant = sum(compliant$compliant == FALSE); num_uncompliant

d <- d.raw %>% 
  left_join(compliant) %>%
  filter(grepl(pattern = "english", 
               language, ignore.case = TRUE), 
         scale != "training1",  
         compliant == TRUE)

# Sample size post compliance check
n.post.compliance = sum(compliant$compliant); n.post.compliance
```

## Cleaning / formatting - output
```{r}
# Preliminary data cleaning
# -------------------------
ms = d %>%
  mutate(stars = as.numeric(as.character(stars))/20) %>%
  group_by(scale, degree, stars) %>%
  summarise(cil = ci.low(judgment), 
            cih = ci.high(judgment), 
            cnt.judgment = sum(judgment),
            speaker.p = mean(judgment)) 
str(ms)
head(ms)
# Output to CSV
# --------------
# write.csv(ms, "../models/L0_e10.csv", row.names=FALSE)
```

# Pragmatic Listener study e6 output

## Read in data
```{r read_data3, message=FALSE, warning=FALSE}
results_path = "../production-results/e6/"
files = dir(results_path, pattern = "*.json")
d.raw = data.frame()

for (f in files) {
  jf = paste(results_path,f,sep="")
  jd = fromJSON(paste(readLines(jf), collapse=""))
  id = data.frame(workerid = jd$WorkerId, 
                   scale = jd$answer$scale,
                   degree = jd$answer$degree,
                   judgment = as.numeric(jd$answer$judgment),
                   language = jd$answer$language)
  d.raw = bind_rows(d.raw, id)
}
```

## Compliance checks

Enforce compliance in e6.

1. training1:high  ->

2. training:low    -> 
```{r}
# Sample size pre compliance check
n.pre.compliance = length(unique(d.raw$workerid)); n.pre.compliance

compliant = d.raw %>% 
  filter(grepl("training", scale)) %>%
  group_by(workerid) %>%
  summarise(compliant = judgment[scale == "training1" & degree == "hi"] > 3 & 
              judgment[scale == "training1" & degree == "low"] < 3)

d = d.raw %>% 
  left_join(compliant) %>%
  filter(grepl(pattern = "english", 
               language, ignore.case = TRUE), 
         scale != "training1", 
         compliant == TRUE)

# Sample size post compliance check
n.post.compliance = length(unique(d$workerid)); n.post.compliance
```

## Cleaning / formatting - output
```{r}
# Preliminary data cleaning
# -------------------------
ms <- d %>%
  group_by(degree, scale) %>%
  mutate(n = n()) %>%
  group_by(degree, scale, judgment) %>%
  summarise(prop = n() / n[1])
str(ms)
head(ms)

# Output to CSV
# --------------
# write.csv(ms, "../models/L1_e6.csv", row.names=FALSE)
```

# Pragmatic Listener study e11 output

## Read in data
```{r read_data4, message=FALSE, warning=FALSE}
results_path = "../production-results/e11/"
files = dir(results_path, pattern = "*.json")
d.raw = data.frame()

for (f in files) {
  jf = paste(results_path,f,sep="")
  jd = fromJSON(paste(readLines(jf), collapse=""))
  id = data.frame(workerid = jd$WorkerId, 
                   scale = jd$answer$data$scale,
                   degree = jd$answer$data$degree,
                   judgment = as.numeric(jd$answer$data$judgment),
                   language = jd$answer$data$language)
  d.raw = bind_rows(d.raw, id)
}
```

## Compliance checks

Enforce compliance in e11.

1. training1:high  -> present "Someone said the food derved a `high` rating". Respondent must choose > 3 stars

2. training:low    -> present "Someone said the food deserved a `low` rating". Respondnet must choose < 3 stars
```{r}
# Sample size pre compliance check
n.pre.compliance = length(unique(d.raw$workerid)); n.pre.compliance

compliant <- d.raw %>% 
   filter(grepl("training", scale)) %>%
   group_by(workerid) %>%
   summarise(compliant = judgment[scale == "training1" & degree == "hi2"] > 3 & 
               judgment[scale == "training1" & degree == "low1"] < 3)

d <- d.raw %>% 
  left_join(compliant) %>%
  filter(grepl(pattern = "english", 
               language, ignore.case = TRUE), 
         scale != "training1", 
         compliant == TRUE)

# Sample size post compliance check
n.post.compliance = length(unique(d$workerid)); n.post.compliance
```

## Cleaning / formatting - output
```{r}
# Preliminary data cleaning
# -------------------------
ms <- d %>%
  group_by(degree, scale) %>%
  mutate(n = n()) %>%
  group_by(degree, scale, judgment) %>%
  summarise(prop = n() / n[1])
# Output to CSV
# --------------
# write.csv(ms, "../models/L1_e11.csv", row.names=FALSE)
```