### 2015-11-03 models/RSA+AmbiguousSemantics.Rmd
populating basic L0 studies information

---
title: "RSA + Scalar implicature / TBD"
author: "BPeloquin/MFrank"
date: "October 29, 2015"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
    toc_depth: 2
---
```{r load_data, results="hide", message=FALSE, warning=FALSE}
rm(list = ls())
setwd("/Users/benpeloquin/Desktop/Projects/scalar_implicature/models")
source("../analysis/useful_dplyr.R")
library(rjson)
library(ggplot2)
```

# Pragmatics / implicature introduction

## Rational Speech-act theory

## Extending RSA to scalar implicature

# Experimental investigations / star rating paradigm

## Literal listener studies

## Can we elicit literal listener semantics for scalar items?

Using the paradigm described in the `star rating paridigm` setting we attempted to elicit literal listener semantics.
```{r}
l0.listener = read.csv("model_data/L0_e8.csv")
str(l0.listener)
head(l0.listener)
```
Data from each literal listener study:

`scale`: The scale we're currently investigating (i.e. `good_excellent`).

`degree`: Valence of the current scalar item (i.e. in the scale `good_excellent` the scalar item `good` has degree `low` and `excellent` has degree `hi`).

`stars`: Number of stars the subject was presented

`cih` and `cil`: lower and upper 95% confidence intervals on the probability average number of positive judgments

`cnt.judgment`: Count of positive responses (`yes`) signalling compatibility between a scalar item and a star rating.

`speaker.p`: Estimated probability of a speaker believing a given scalar items is compatible with a presented star rating.

### Literal Listener study - two scalar items

We can approximate the idea of `literal listener` semantics as a distribution over star ratings.
```{r}
qplot(x = stars, y = speaker.p, col = degree,
      geom = "line", stat = "identity", position = "dodge",
      main = "Scalar compatibility with star ratings \n Experiment L0_e8",
      ylab = "P(Word | Rating)",
      data = l0.listener) +
  geom_linerange(aes(ymin = speaker.p - cil, 
                     ymax = speaker.p + cih)) + 
  facet_wrap(~scale)
```

### Literal Listener study - Full scale from empirically derived alternatives
```{r}
l0.listener2 = read.csv("model_data/L0_e10.csv")
str(l0.listener2)
head(l0.listener2)
```

`literal listener` semantics with full distribution over ratings.
```{r}
qplot(x = stars, y = speaker.p, col = degree,
      geom = "line", stat = "identity", position = "dodge", 
      main = "Scalar compatibility with star ratings \n Experiment L0_e10",
      ylab = "P(Word | Rating)",
      data = l0.listener2) + 
  geom_linerange(aes(ymin = speaker.p - cil, 
                     ymax = speaker.p + cih)) + 
  facet_wrap(~scale)
```

Have the distributions changed between studies (for the shared items)?
```{r}
subset.l2 = which((l0.listener2[, "scale"] != "some_all" &
        (l0.listener2[, "degree"] == "hi2" | l0.listener2[, "degree"] == "hi1")) |
       (l0.listener2[, "scale"] == "some_all" &
          (l0.listener2[, "degree"] == "hi2" | l0.listener2[, "degree"] == "low1")))

l0.listener2.match = l0.listener2[subset.l2,]
l0.listener2.match$degree = ifelse(l0.listener2.match$degree == "hi2", "hi", "low")

qplot(x = stars, y = speaker.p, col = degree,
      geom = "line", stat = "identity", position = "dodge", 
      main = "Scalar compatibility with star ratings \n Experiment L0_e10",
      ylab = "P(Word | Rating)",
      data = l0.listener2.match) + 
  geom_linerange(aes(ymin = speaker.p - cil, 
                     ymax = speaker.p + cih)) + 
  facet_wrap(~scale)
```

### Description

### Results

## Pragmatic listener studies

### Description

### Results

# Formalizing Grice / RSA / Bayesian model

## Current model description

## Implementation

# Connecting experiments and model

## Baseline model performance

## Miniumum alternatives model

## Eliciting alternatives experiment

### Description

### Results

## Full model (alternatives) performance

#Conclusions