### 2015-11-29 models/Model_Comparison.Rmd
starting write-up for model comparisons

---
title: "Model Comparisons"
author: "BPeloquin"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
    toc_depth: 2
---

```{r include=FALSE}
rm(list=ls())

library(ggplot2)
library(knitr)
library(rjson)
library(dplyr)
library(tidyr)
library(gridExtra)
```

# RSA functionality
```{r}
# speaker informativity
# ---------------------
speaker.inf = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd = function(rating, degree, m, alpha) {
  numerator = speaker.inf(m[rating, degree], alpha)
  normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
  return(numerator / normalize)
}
# non-normalized posterior
# -----------------------
nn.post = function(rating, degree, m, alpha, useprior) {
  prior = priors[rating, "prior.p"]
  return(speaker.lhd(rating, degree, m, alpha) * prior)
}
# normalized posterior
# --------------------
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  normalize = sum(unlist(sapply(seq(1, 5), function(i){nn.post(i, degree, m, alpha, useprior)})))
  return(nn / normalize)
}
```

# Run model functionality
```{r}
# run.partial()
# ------------
# Run RSA with model1 (entailment) and model2 (entailment + generic)
run.partial = function(d, alpha=1, useprior=F, usenone=F, normalize=F) {
  if (normalize) {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      mutate(hi = hi / sum(hi),
             low = low / sum(low)) %>%
      select(hi, low)
  } else {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      select(hi, low)
  }

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  }

  d$pred = round(as.numeric(mapply(norm.post, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
}

# run.full()
# ------------
# Run RSA with model3 (full) with alternatives
run.full = function(d, alpha=1, useprior=F, usenone=F, addMid=F, normalize=F) {
  # alpha = scales.entropy[scales.entropy$scale==d$scale[1], ]$Entropy
  if (addMid) {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               mid = mid / sum(mid),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, mid, low1, low2)    
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, mid, low1, low2)    
    }
  } else {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, low1, low2)
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, low1, low2)
    }
  }

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  }

  d$pred = round(as.numeric(mapply(norm.post, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
}
```

## Tune hyperparams
```{r}
# tune.alhpa()
# ------------
# d        --> data
# alphas   --> range of alphas to test
# type     --> full or partial model
# useprior --> use uniform prior
# usenone  --> use gener None
tune.alpha = function(d, alphas = seq(from=1, to=10),
                      type="partial", useprior = T,
                      usenone=F, compare.data=NULL, addMid = F, normalize=F) {
  # Tune best alphas
  fit = sapply(alphas, FUN=function(n) {
    if (type == "partial") {
      md = d %>%
        do(run.partial(., alpha=n, useprior=useprior, usenone=usenone, normalize=normalize))
    } else {
      md = d %>%
        do(run.full(., alpha=n, useprior=useprior, usenone=usenone, addMid=addMid, normalize=normalize))
    }
    # Toggle fit to e6 data
    if (!is.null(compare.data)) {

      # If we're using e11 data with multiple scalars
      if ("hi1" %in% md$degree) {
        matched.items = which((md[, "scale"] != "some_all" &
                                 (md[, "degree"] == "hi2" | md[, "degree"] == "hi1")) |
                                (md[, "scale"] == "some_all" &
                                   (md[, "degree"] == "hi1" | md[, "degree"] == "low1")))
        md = md[matched.items, ]
        md$degree = ifelse(md$degree == "hi1", "hi", "low")
        stopifnot("listener.p" %in% colnames(compare.data))
      }
      md$listener.p = compare.data$listener.p
    }

    # MSE
    return(mean((md$pred - md$listener.p)^2))
  })  
  # get lowest MSE
  best.alpha = which(fit == min(fit))
  return(best.alpha)
}

# Uniform priors
unif.priors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
emp.priors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
priors = unif.priors

scales.entropy = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/scales_entropy.csv")
```

# Model Comparisons

### Data for entailment models
```{r entailment_models_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
# Combine speaker / listener
data.partial = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

### Data for full model - no alternatives
```{r full_model_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv")
data.full = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

### Data for full model - alternatives
```{r}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10a.csv")
data.full.extras = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

## Match items between studies (full and partial)
```{r}
matched.items = which((data.full[, "scale"] != "some_all" &
        (data.full[, "degree"] == "hi2" | data.full[, "degree"] == "hi1")) |
       (data.full[, "scale"] == "some_all" &
          (data.full[, "degree"] == "hi1" | data.full[, "degree"] == "low1")))
matched.items.extras = which((data.full.extras[, "scale"] != "some_all" &
        (data.full.extras[, "degree"] == "hi2" | data.full.extras[, "degree"] == "hi1")) |
       (data.full.extras[, "scale"] == "some_all" &
          (data.full.extras[, "degree"] == "hi1" | data.full.extras[, "degree"] == "low1")))
```

## Run comparisons
```{r run_comparisons, message=FALSE}
# Save performance output
performance.output = data.frame(model=rep(NA, 16),
                                cor.e6=rep(NA, 16),
                                cor.e11=rep(NA, 16),
                                normalized=rep(NA, 16))
alphas = rep(NA, 16)
data.full.transfer = data.full[matched.items, ]
data.full.transfer$degree =
  ifelse(data.full.transfer$degree == "hi1", "hi", "low")
# Model 1 - Entailment only models
# --------------------------------
# 1.a)
# Entailment only
# Normalized = T
# alpha tuning = F
alphas[1] = 1
m1.a = data.partial %>%
  do(run.partial(., alpha = alphas[1], useprior=T, usenone=F, normalize=T))
# Store output and add e11
m1.a = cbind(m1.a, data.full.transfer$listener.p)
colnames(m1.a)[8] = "e11.listener.p"
performance.output[1, ] = c("m1.a", round(cor(m1.a$listener.p, m1.a$pred), 5),
                            round(cor(m1.a$e11.listener.p, m1.a$pred), 5), T)
# Store plot
m1.a.plot = qplot(stars, listener.p, col=degree, 
      data=m1.a,
      main=paste("m1.a\nalpha: ", alphas[1],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[1, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 1.b)
# Entailment only
# Normalized = F
# alpha tune = F
alphas[2] = 1
m1.b = data.partial %>%
  do(run.partial(., alpha = alphas[1], useprior=T, usenone=F, normalize=F))
# Store output
m1.b = cbind(m1.b, data.full.transfer$listener.p)
colnames(m1.b)[8] = "e11.listener.p"
performance.output[2, ] = c("m1.b", round(cor(m1.b$listener.p, m1.b$pred), 5),
                            round(cor(m1.b$e11.listener.p, m1.b$pred), 5), F)
# Store plot
m1.b.plot = qplot(stars, listener.p, col=degree, 
      data=m1.b,
      main=paste("m1.b\nalpha: ", alphas[2],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[2, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 1.c)
# Entailment only
# Normalized = T
# alpha tune = T
alphas[3] = tune.alpha(data.partial, normalize=T, compare.data=data.full.transfer)
m1.c = data.partial %>%
  do(run.partial(., alpha = alphas[3], useprior=T, usenone=F, normalize=T))
# Store output
m1.c = cbind(m1.c, data.full.transfer$listener.p)
colnames(m1.c)[8] = "e11.listener.p"
performance.output[3, ] = c("m1.c", round(cor(m1.c$listener.p, m1.c$pred), 5),
                            round(cor(m1.c$e11.listener.p, m1.c$pred), 5), T)
# Store plot
m1.c.plot = qplot(stars, listener.p, col=degree, 
      data=m1.c,
      main=paste("m1.c\nalpha: ", alphas[3],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[3, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 1.d)
# Entailment only
# Normalized = F
# alpha tune = T
alphas[4] = tune.alpha(data.partial, normalize=F)
m1.d = data.partial %>%
  do(run.partial(., alpha = alphas[4], useprior=T, usenone=F, normalize=F))
# Store output
m1.d = cbind(m1.d, data.full.transfer$listener.p)
colnames(m1.d)[8] = "e11.listener.p"
performance.output[4, ] = c("m1.d", round(cor(m1.d$listener.p, m1.d$pred), 5),
                            round(cor(m1.d$e11.listener.p, m1.d$pred), 5), normalized=F)
# Store plot
m1.d.plot = qplot(stars, listener.p, col=degree, 
      data=m1.d,
      main=paste("m1.d\nalpha: ", alphas[4],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[4, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# Model 2 - Entailment + generic none
# --------------------------------
# 2.a)
# Entailment + generic None
# Normalized = T
# alpha tuning = F
alphas[5] = 1
m2.a = data.partial %>%
  do(run.partial(., alpha = alphas[5], useprior=T, usenone=T, normalize=T))
# Store output
m2.a = cbind(m2.a, data.full.transfer$listener.p)
colnames(m2.a)[8] = "e11.listener.p"
performance.output[5, ] = c("m2.a", round(cor(m2.a$listener.p, m2.a$pred), 5),
                            round(cor(m2.a$e11.listener.p, m2.a$pred), 5), T)
# Store plot
m2.a.plot = qplot(stars, listener.p, col=degree, 
      data=m2.a,
      main=paste("m2.a\nalpha: ", alphas[5],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[5, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 2.b)
# Entailment + generic None
# Normalized = F
# alpha tune = F
alphas[6] = 1
m2.b = data.partial %>%
  do(run.partial(., alpha = alphas[6], useprior=T, usenone=T, normalize=F))
# Store output
m2.b = cbind(m2.b, data.full.transfer$listener.p)
colnames(m2.b)[8] = "e11.listener.p"
performance.output[6, ] = c("m2.b", round(cor(m2.b$listener.p, m2.b$pred), 5),
                            round(cor(m2.b$e11.listener.p, m2.b$pred), 5), F)
# Store plot
m2.b.plot = qplot(stars, listener.p, col=degree, 
      data=m2.b,
      main=paste("m2.b\nalpha: ", alphas[6],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[6, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 2.c)
# Entailment + generic None
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.full.transfer for e11 or NULL for e6
alphas[7] = tune.alpha(data.partial, normalize=T, compare.data=NULL)
m2.c = data.partial %>%
  do(run.partial(., alpha = alphas[7], useprior=T, usenone=T, normalize=T))
# Store output
m2.c = cbind(m2.c, data.full.transfer$listener.p)
colnames(m2.c)[8] = "e11.listener.p"
performance.output[7, ] = c("m2.c", round(cor(m2.c$listener.p, m2.c$pred), 5),
                            round(cor(m2.c$e11.listener.p, m2.c$pred), 5), T)
# Store plot
m2.c.plot = qplot(stars, listener.p, col=degree, 
      data=m2.c,
      main=paste("m2.c\nalpha: ", alphas[7],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[7, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 2.d)
# Entailment + generic None
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.full.transfer for e11 or NULL for e6
alphas[8] = tune.alpha(data.partial, normalize=F)
m2.d = data.partial %>%
  do(run.partial(., alpha = alphas[8], useprior=T, usenone=T, normalize=F))
# Store output
m2.d = cbind(m2.d, data.full.transfer$listener.p)
colnames(m2.d)[8] = "e11.listener.p"
performance.output[8, ] = c("m2.d", round(cor(m2.d$listener.p, m2.d$pred), 5),
                            round(cor(m2.d$e11.listener.p, m2.d$pred), 5), F)
# Store plot
m2.d.plot = qplot(stars, listener.p, col=degree, 
      data=m2.d,
      main=paste("m2.d\nalpha: ", alphas[8],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[8, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# Model 3 - Empirical alternatives
# --------------------------------
# 3.a)
# Emp alts
# Normalized = T
# alpha tuning = F
alphas[9] = 1
m3.a = data.full %>%
  do(run.full(., alpha = alphas[9], addMid=F, normalize=T))
# Match with e6 w/ checks
m3.a.matched = m3.a[matched.items, ]
m3.a.matched$degree = ifelse(m3.a.matched$degree == "hi1", "hi", "low")
m3.a.matched = cbind(m3.a.matched, data.partial$listener.p)
all(m3.a.matched$scale == data.partial$scale & m3.a.matched$degree == data.partial$degree)
colnames(m3.a.matched)[length(colnames(m3.a.matched))] = "e6.listener.p"
# Store output
performance.output[9, ] = c("m3.a", round(cor(m3.a.matched$e6.listener.p, m3.a.matched$pred), 5),
                            round(cor(m3.a.matched$listener.p, m3.a.matched$pred), 5), T)
# Store plot
m3.a.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.a.matched,
      main=paste("m3.a\nalpha: ", alphas[9],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[9, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 3.b)
# Emp alts
# Entailment + generic None
# Normalized = F
# alpha tune = F
alphas[10] = 1
m3.b = data.full %>%
  do(run.full(., alpha = alphas[10], addMid=F, normalize=F))
# Match with e6 w/ checks
m3.b.matched = m3.b[matched.items, ]
m3.b.matched$degree = ifelse(m3.b.matched$degree == "hi1", "hi", "low")
m3.b.matched = cbind(m3.b.matched, data.partial$listener.p)
all(m3.b.matched$scale == data.partial$scale & m3.b.matched$degree == data.partial$degree)
colnames(m3.b.matched)[length(colnames(m3.b.matched))] = "e6.listener.p"
# Store output
performance.output[10, ] = c("m3.b", round(cor(m3.b.matched$e6.listener.p, m3.b.matched$pred), 5),
                             round(cor(m3.b.matched$listener.p, m3.b.matched$pred), 5), F)
# Store plot
m3.b.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.b.matched,
      main=paste("m3.b\nalpha: ", alphas[10],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[10, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 3.c)
# Emp alts
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.partial for e6 or NULL for e11
alphas[11] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=T)
m3.c = data.full %>%
  do(run.full(., alpha = alphas[11], addMid=F, normalize=T))
# Match with e6 w/ checks
m3.c.matched = m3.c[matched.items, ]
m3.c.matched$degree = ifelse(m3.c.matched$degree == "hi1", "hi", "low")
m3.c.matched = cbind(m3.c.matched, data.partial$listener.p)
all(m3.c.matched$scale == data.partial$scale & m3.c.matched$degree == data.partial$degree)
colnames(m3.c.matched)[length(colnames(m3.c.matched))] = "e6.listener.p"
# Store output
performance.output[11, ] = c("m3.c", round(cor(m3.c.matched$e6.listener.p, m3.c.matched$pred), 5),
                             round(cor(m3.c.matched$listener.p, m3.c.matched$pred), 5), T)
# Store plot
m3.c.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.c.matched,
      main=paste("m3.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[11, 2]),
      ylab="Posterior p(rating | word)") +
      ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 3.d)
# Emp alts
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[12] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=F)
m3.d = data.full %>%
  do(run.full(., alpha = alphas[12], addMid=F, normalize=F))
# Match with e6 w/ checks
m3.d.matched = m3.d[matched.items, ]
m3.d.matched$degree = ifelse(m3.d.matched$degree == "hi1", "hi", "low")
m3.d.matched = cbind(m3.d.matched, data.partial$listener.p)
all(m3.d.matched$scale == data.partial$scale & m3.d.matched$degree == data.partial$degree)
colnames(m3.d.matched)[length(colnames(m3.d.matched))] = "e6.listener.p"
# Store output
performance.output[12, ] = c("m3.d", round(cor(m3.d.matched$e6.listener.p, m3.d.matched$pred), 5),
                             round(cor(m3.d.matched$listener.p, m3.d.matched$pred), 5), F)
# Store plot
m3.d.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.d.matched,
      main=paste("m3.d\nalpha: ", alphas[12],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[12, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# Model 4 - Empirical alternatives + extras
# -----------------------------------------
# 4.a)
# Emp alts + extras
# Normalized = T
# alpha tuning = F
alphas[13] = 1
m4.a = data.full.extras %>%
  do(run.full(., alpha = alphas[13], addMid=T, normalize=T))
# Match with e6 w/ checks
m4.a.matched = m4.a[matched.items.extras, ]
m4.a.matched$degree = ifelse(m4.a.matched$degree == "hi1", "hi", "low")
m4.a.matched = cbind(m4.a.matched, data.partial$listener.p)
all(m4.a.matched$scale == data.partial$scale & m4.a.matched$degree == data.partial$degree)
colnames(m4.a.matched)[length(colnames(m4.a.matched))] = "e6.listener.p"
# Store output
performance.output[13, ] = c("m4.a", round(cor(m4.a.matched$e6.listener.p, m4.a.matched$pred), 5), 
                             round(cor(m4.a.matched$listener.p, m4.a.matched$pred), 5), T)
# Store plot
m4.a.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.a.matched,
      main=paste("m4.a\nalpha: ", alphas[13],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[13, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.b)
# Emp alts + extras
# Normalized = F
# alpha tune = F
alphas[14] = 1
m4.b = data.full.extras %>%
  do(run.full(., alpha = alphas[14], addMid=T, normalize=F))
# Match with e6 w/ checks
m4.b.matched = m4.b[matched.items.extras, ]
m4.b.matched$degree = ifelse(m4.b.matched$degree == "hi1", "hi", "low")
m4.b.matched = cbind(m4.b.matched, data.partial$listener.p)
all(m4.b.matched$scale == data.partial$scale & m4.b.matched$degree == data.partial$degree)
colnames(m4.b.matched)[length(colnames(m4.b.matched))] = "e6.listener.p"
# Store output
performance.output[14, ] = c("m4.b", round(cor(m4.b.matched$e6.listener.p, m4.b.matched$pred), 5), 
                             round(cor(m4.b.matched$listener.p, m4.b.matched$pred), 5), F)
# Store plot
m4.b.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.b.matched,
      main=paste("m4.b\nalpha: ", alphas[14],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[14, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.c)
# Emp alts + extras
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[15] = tune.alpha(data.full.extras, type = "full", compare.data=data.partial, addMid=T, normalize=T)
m4.c = data.full.extras %>%
  do(run.full(., alpha = alphas[15], addMid=T, normalize=T))
# Match with e6 w/ checks
m4.c.matched = m4.c[matched.items.extras, ]
m4.c.matched$degree = ifelse(m4.c.matched$degree == "hi1", "hi", "low")
m4.c.matched = cbind(m4.c.matched, data.partial$listener.p)
all(m4.c.matched$scale == data.partial$scale & m4.c.matched$degree == data.partial$degree)
colnames(m4.c.matched)[length(colnames(m4.c.matched))] = "e6.listener.p"
# Store output (e11 and e6)
performance.output[15, ] = c("m4.c", round(cor(m4.c.matched$e6.listener.p, m4.c.matched$pred), 5),
                             round(cor(m4.c.matched$listener.p, m4.c.matched$pred), 5), T)

# Store plot
m4.c.plot = qplot(stars, listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.d)
# Emp alts + extras
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[16] = tune.alpha(data.full.extras, type = "full", compare.data=data.partial, addMid=T, normalize=F)
m4.d = data.full.extras %>%
  do(run.full(., alpha = alphas[16], addMid=T, normalize=F))
# Match with e6 w/ checks
m4.d.matched = m4.d[matched.items.extras, ]
m4.d.matched$degree = ifelse(m4.d.matched$degree == "hi1", "hi", "low")
m4.d.matched = cbind(m4.d.matched, data.partial$listener.p)
all(m4.d.matched$scale == data.partial$scale & m4.d.matched$degree == data.partial$degree)
colnames(m4.d.matched)[length(colnames(m4.d.matched))] = "e6.listener.p"
# Store output
performance.output[16, ] = c("m4.d", round(cor(m4.d.matched$e6.listener.p, m4.d.matched$pred), 5),
                             round(cor(m4.d.matched$listener.p, m4.d.matched$pred), 5), F)
# Store plot
m4.d.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.d.matched,
      main=paste("m4.d\nalpha: ", alphas[16],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[16, 2]),
      ylab="Posterior p(rating | word)") +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

# Plots

## Interpreting the output table:

`m1:` Entailment only models (i.e. "some_all" scale only includes "some", "all")

`m2:` Entailment + generic None models (i.e. "some_all" scale includes "some", "all", "none")

`m3:` Full model with empirical alternatives (i.e. "some_all" scale includes "some", "all", "most", "none")

`m4:` Entailment + empirical alternatives + additional fitted scalar (i.e. "some_all" scale includes "some", "all", "most", "none", "mid")

`a:` Normalized literal listener values before running RSA, no alpha tuning

`b:` No normalization for literal listener values before running RSA, no alpha tuning

`c:` Normalized literal listener values before running RSA, alpha tuning

`d:` No normalization for literal listener values before running RSA, alpha tuning

Initial thoughts:

We see fairly steady model improvement with the addition of salient alternatives. Let's focus on the normalized family of models (1a, 1c, ... , 4a, 4c) - I explain why this might be a good idea below, but I'm not totally convinced. Probably the most important comparisons to pay attention to is the jump in correlation values when we first introduce empirically derived alternatives (models m3). Comparing m2 models (entailment + generic None) and m3 models (empirically derived alternatives) we see jumps of +0.23 for r values with alpha tuning. See the plots below:
```{r}
m2.c.plot
m3.c.plot
```
Importantly, we also see improvements with the addition of 'fitted' alternatives in m4 see the plots below:
```{r}
m3.c.plot
m4.c.plot
```

Additional notes:

1) In general model performance is greater (correlation) for e11 data. This may be due in part to less extreme literal listener evaluations in e11 (see RSA+AmbiguousSemantics for a more detailed analysis for literal listener experiment differences). Our best model performance occurs with e11 and the model is m4.c, correlation = `r round(performance.outpu[15, 3], 5)`.

2) Non-normalized models seem to be more sensitive to alpha tuning as the number of alternatives increases (see m4.d and m3.d with alpha = 8). I'm inclined to stick with normalized model results because of this - we're interested in whether we see model improvement from the addition of salient alternatives - not from tuning hyperparameters. We tend to see more model improvement with alternatives when we're looking at normalized models (see `m3.c` vs `m4.c`).

3) We may want to explore the the idea of unique alphas for a given scale (would need to change how we've written RSA), for example using scale entropy measures.
```{r overall_perf}
performance.output = cbind(performance.output, alphas)
grid.table(performance.output)

m1.a.plot
m1.b.plot
m1.c.plot
m1.d.plot

m2.a.plot
m2.b.plot
m2.c.plot
m2.d.plot

m3.a.plot
m3.b.plot
m3.c.plot
m3.d.plot

m4.a.plot
m4.b.plot
m4.c.plot
m4.d.plot
```

# Other exploration
```{r}
# bad.predictions = m3.fit.matched$pred > 0 &
#                   m3.fit.matched$e6.listener.p == 0 &
#                   (m3.fit.matched$pred - m3.fit.matched$e6.listener.p > 0.05)
```