### 2015-12-01 models/Model_additionalAlts.Rmd
modiefied gradient descent tracking and plotting

---
title: "Model_additionalAlts"
author: "BPeloquin"
date: "November 25, 2015"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
    toc_depth: 3
---
```{r include=FALSE}
rm(list=ls())

library(ggplot2)
library(knitr)
library(rjson)
library(dplyr)
library(tidyr)
library(gridExtra)
```

# RSA functionality
```{r}
# speaker informativity
# ---------------------
speaker.inf = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd = function(rating, degree, m, alpha) {
  numerator = speaker.inf(m[rating, degree], alpha)
  normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
  return(numerator / normalize)
}
# non-normalized posterior
# -----------------------
nn.post = function(rating, degree, m, alpha, useprior) {
  prior = priors[rating, "prior.p"]
  return(speaker.lhd(rating, degree, m, alpha) * prior)
}
# normalized posterior
# --------------------
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  normalize = sum(unlist(sapply(seq(1, 5), function(i){nn.post(i, degree, m, alpha, useprior)})))
  return(nn / normalize)
}
```

# Run model functionality
```{r}
# run.partial()
# ------------
# Run RSA with model1 (entailment) and model2 (entailment + generic)
run.partial = function(d, alpha=1, useprior=F, usenone=F, normalize=F) {
  if (normalize) {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      mutate(hi = hi / sum(hi),
             low = low / sum(low)) %>%
      select(hi, low)
  } else {
    mat = d %>%
      select(stars, degree, speaker.p) %>%
      spread(degree, speaker.p) %>%
      select(hi, low)
  }

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  }

  d$pred = round(as.numeric(mapply(norm.post, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
}

# run.full()
# ------------
# Run RSA with model3 (full) with alternatives
run.full = function(d, alpha=1, useprior=F, usenone=F, addMid=F, normalize=F) {
  if (addMid) {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               mid = mid / sum(mid),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, mid, low1, low2)    
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, mid, low1, low2)    
    }
  } else {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, low1, low2)
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, low1, low2)
    }
  }

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  }

  d$pred = round(as.numeric(mapply(norm.post, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
}
```

## Tune hyperparams
```{r}
# tune.alhpa()
# ------------
# d        --> data
# alphas   --> range of alphas to test
# type     --> full or partial model
# useprior --> use uniform prior
# usenone  --> use gener None
tune.alpha = function(d, alphas = seq(from=1, to=10),
                      type="partial", useprior = T,
                      usenone=F, compare.data=NULL, addMid = F, normalize=F) {
  # Tune best alphas
  fit = sapply(alphas, FUN=function(n) {
    if (type == "partial") {
      md = d %>%
        do(run.partial(., alpha=n, useprior=useprior, usenone=usenone, normalize=normalize))
    } else {
      md = d %>%
        do(run.full(., alpha=n, useprior=useprior, usenone=usenone, addMid=addMid, normalize=normalize))
    }
    # Fit to e6 data
    if (!is.null(compare.data)) {
      matched.items = which((md[, "scale"] != "some_all" &
                               (md[, "degree"] == "hi2" | md[, "degree"] == "hi1")) |
                              (md[, "scale"] == "some_all" &
                                 (md[, "degree"] == "hi1" | md[, "degree"] == "low1")))
      md = md[matched.items, ]
      md$degree = ifelse(md$degree == "hi1", "hi", "low")
      stopifnot("listener.p" %in% colnames(compare.data))
      md$listener.p = compare.data$listener.p
    }

    # MSE
    return(mean((md$pred - md$listener.p)^2))
  })  
  # get lowest MSE
  best.alpha = which(fit == min(fit))
  return(best.alpha)
}

# Uniform priors
unif.priors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
emp.priors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
priors = unif.priors

scales.entropy = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/scales_entropy.csv")
```

# Model Comparisons

### Data for entailment models
```{r entailment_models_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")
# Combine speaker / listener
data.partial = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

### Data for full model - no alternatives
```{r full_model_data, message=FALSE, warning=FALSE}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
listener = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e11.csv")
data.full = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)

data.full$entropy = scales.entropy[data.full$scale, ]$Entropy
```

### Data for full model - alternatives
```{r}
speaker = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10a.csv")
data.full.extras = left_join(speaker, listener) %>%
  left_join(priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, listener.p, prior.p) %>%
  mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p)) %>%
  group_by(scale)
```

## Match items between studies (full and partial)
```{r}
matched.items = which((data.full[, "scale"] != "some_all" &
        (data.full[, "degree"] == "hi2" | data.full[, "degree"] == "hi1")) |
       (data.full[, "scale"] == "some_all" &
          (data.full[, "degree"] == "hi1" | data.full[, "degree"] == "low1")))
matched.items.extras = which((data.full.extras[, "scale"] != "some_all" &
        (data.full.extras[, "degree"] == "hi2" | data.full.extras[, "degree"] == "hi1")) |
       (data.full.extras[, "scale"] == "some_all" &
          (data.full.extras[, "degree"] == "hi1" | data.full.extras[, "degree"] == "low1")))
```

# Fit extra scalar literal listener distributions
## modified gradient descent / MLE fn()
```{r}
# normalize()
# -----------
# normalize a generic vector
normalize = function(w) { return (w / sum(w))}

# fit_distr()
# -----------
# target        --> specific scalar family we're interested in (i.e. "some_all")
# numIters      --> number of iterations to run
# eta           --> step size
# alpha         --> alpha level for RSA?
# normalize     --> normalize in RSA?
fit_distr = function(target, numIters = 100, eta=0.01, normalize=T,
                     alpha=1, verbose=F, data.record=NULL) {

  if (verbose) { print(paste("Currently fitting", target, "...")) }
  # Initialize to uniform
  w = rep(0.1, 5)
  # Focus on scalar 'target'
  scalar.focused = data.full.extras[data.full.extras$scale == target, ]

  # For numIters iterations...
  for (i in 1:numIters) {
    prev.w = w
    scalar.focused[scalar.focused$degree == "mid", "speaker.p"] = w
    m = scalar.focused %>%
      do(run.full(., alpha = alpha, addMid=T, normalize=normalize))
    w[1] = w[1] - eta * (sum(m[m$stars == 1, ]$listener.p - m[m$stars == 1, ]$pred))
    w[2] = w[2] - eta * (sum(m[m$stars == 2, ]$listener.p - m[m$stars == 2, ]$pred))
    w[3] = w[3] - eta * (sum(m[m$stars == 3, ]$listener.p - m[m$stars == 3, ]$pred))
    w[4] = w[4] - eta * (sum(m[m$stars == 4, ]$listener.p - m[m$stars == 4, ]$pred))
    w[5] = w[5] - eta * (sum(m[m$stars == 5, ]$listener.p - m[m$stars == 5, ]$pred))
    w = ifelse(w <= 0, 0, w)
    w = ifelse(w >= 1, 1, w)
    w = normalize(w)

    # Record convergence
    if (!is.null(data.record)) {
      data.record = rbind(data.record, w)
    }

    # Check if we've converged
    if (sum(abs(prev.w - w)) < 0.00001) {
      if(verbose) { print("Convergence") }
      return(list(w, data.record))
    }

    if(verbose) {print(list(i, w))}
  }
  return(list(w, data.record))
}
```

# Fit extra scalar literal listener distributions
```{r}
# Normalize in RSA?
norm = T
# Our scalars
scalars = c("good_excellent",
            "liked_loved",
            "palatable_delicious",
            "memorable_unforgettable",
            "some_all")
fitted.distr = data.frame(matrix(rep(NA, 25), ncol=5))
colnames(fitted.distr) = scalars

# Fit each of our scalars / track fitting
track.data = data.frame(matrix(ncol=6))
for (i in 1:length(scalars)) {
  # record fitting
  data.record = data.frame(matrix(rep(NA, 5), ncol=5))
  # output is optimal distr [[1]] and [[2]] record of fitting
  out = fit_distr(scalars[i], numIters=20, eta=0.01,
                  verbose=F, normalize=norm, alpha=4, data.record=data.record)

  # get fitted scalar distr
  fitted.distr[, i] = out[[1]]
  # get tracking data
  tracking = as.data.frame(out[[2]][-1,]) %>%
    mutate(X6 = scalars[i])
  track.data = rbind(track.data, tracking)
}
```

```{r}
# Format track data for display
track.data = track.data[-1, ]
track.data$step = rep(seq(1, 20), 5)
track.data = track.data %>% 
  gather(stars, value, X1:X5)
```

## Run distribution fitting
```{r}
for (scale in scalars) {
  data.full.extras[data.full.extras$scale == scale &
                     data.full.extras$degree == "mid", ]$speaker.p =
    normalize(fitted.distr[ , scale])
}
m = data.full.extras %>%
  do(run.full(., alpha = 4, addMid=T, normalize=norm))
cor(m[matched.items.extras, ]$listener.p, m[matched.items.extras, ]$pred)
```

## Plot distr
```{r}
# Normalize speaker.p for plotting
for (scale in scalars) {
  degrees = unique(data.full.extras$degree)
  for (degree in degrees) {
    data.full.extras[data.full.extras$scale == scale &
                       data.full.extras$degree == degree, ]$speaker.p = 
        normalize(data.full.extras[data.full.extras$scale == scale &
                                     data.full.extras$degree == degree, ]$speaker.p)
  }
}

# Differentiate tuned 'mid' scalar
data.full.extras$type = ifelse(data.full.extras$degree == "mid", "red", "blue")
qplot(x = stars, y = speaker.p, colour = type,
      geom = "line", stat = "identity", position = "dodge", 
      main = "Literal listener semantics\nEmpirical estimates and tuned extra 'mid'",
      ylab = "P_{L0}(Word | Rating) - Normalized",
      data = data.full.extras) + 
  facet_wrap(~scale + degree)

# Tracking of gradient descent tuning
ggplot(data=track.data, aes(x=step, y=value, col=stars)) +
  geom_line() +
  facet_wrap(~X6)
```