### 2016-01-19 models/RSA_alternativeCostFormulation.Rmd
fns() for model runs - entailment / neg alts / full

---
title: "alternativeRSA"
author: "Ben Peloquin"
date: "January 13, 2016"
output: html_document
---

```{r}
rm(list = ls())
library(ggplot2)
library(knitr)
library(rjson)
library(dplyr)
library(tidyr)
library(gridExtra)
```

Old RSA formulation
```{r}
# speaker informativity
# ---------------------
speaker.inf = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd = function(rating, degree, m, alpha) {
  numerator = speaker.inf(m[rating, degree], alpha)
  normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
  return(numerator / normalize)
}
# non-normalized posterior
# -----------------------
nn.post = function(rating, degree, m, alpha, useprior) {
  prior = priors[rating, "prior.p"]
  return(speaker.lhd(rating, degree, m, alpha) * prior)
}
# normalized posterior
# --------------------
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  normalize = sum(unlist(sapply(seq(1, 5),
                                function(i){nn.post(i, degree, m, alpha, useprior)})))
  return(nn / normalize)
}
```

New Prior incorporation
```{r}
speaker.inf3 = function(d, alpha, cost = 0) {
  exp(-1 * alpha * (-log(d) - cost))
  # exp(alpha * (log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd3 = function(currRating, currDegree, m, alpha, costs, allDegrees) {
  currCost <- subset(costs, degree == currDegree)$cost
  numerator = speaker.inf3(m[[currDegree]][currRating], alpha, currCost)
  # normalize over other words
  normalize <- sum(unlist(sapply(seq(allDegrees), FUN = 
                            function(i) {
                              speaker.inf(m[[allDegrees[i]]][currRating], alpha,
                                          subset(costs, degree == allDegrees[i])$cost)}
                          )))
  return(numerator / normalize)
  # return(numerator)
}
# old !!!!
# speaker.lhd = function(rating, degree, m, alpha) {
#   numerator = speaker.inf(m[rating, degree], alpha)
#   normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
#   return(numerator / normalize)
# }

# non-normalized posterior
# -----------------------
nn.post3 = function(currRating, currDegree, m, alpha, costs, priors, allDegrees) {
  # prior <- priors[rating, "prior.p"]
  # prior <- costs[degree]
  currPrior <- subset(priors, stars == currRating)$prior.p
  return(speaker.lhd3(currRating, currDegree, m, alpha, costs, allDegrees) * currPrior)
}
# normalized posterior
# --------------------
norm.post3 = function(currRating, currDegree, m, alpha, costs, priors, allDegrees) {
  nn <- nn.post3(currRating, currDegree, m, alpha, costs, priors, allDegrees)
  normalize <- sum(unlist(sapply(seq(1, 5),
                                function(i) {
                                  nn.post3(i, currDegree, m, alpha, costs, priors, allDegrees)}
                                )))
  return(nn / normalize)
}

normedVec <- function(vec) {
  vec / sum(vec)
}
```

Run RSA - five item scales
```{r}
computePosteriors <- c(norm.post, norm.post3)
# runFull() 
# ---------
# data       => contains scale | degree | stars | speaker.p | prior.p | cost
# calcFN     => without cost = 1, with cost = 2
# alpha      => set alpha level
runFull <- function(data, calcFn = 2, alpha = 1, numAlts = 5) {
  results <-
    data.frame(scale = c(),
               stars = c(),
               degree = c(),
               preds = c())

  scales <- unique(data$scale)
  for (s in scales) {
      if (numAlts == 5) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 mid = mid / sum(mid),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, mid, low1, low2)
      } else if (numAlts == 4) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi1 = hi1 / sum(hi1),
                 hi2 = hi2 / sum(hi2),
                 low1 = low1 / sum(low1),
                 low2 = low2 / sum(low2)) %>%
          select(hi1, hi2, low1, low2)
      } else if (numAlts == 2) {
        mat <- subset(data, scale  == s) %>%
          select(stars, degree, speaker.p) %>%
          spread(degree, speaker.p) %>%
          mutate(hi = hi / sum(hi),
                 low = low / sum(low)) %>%
          select(hi, low)
      }
    stars <- subset(data, scale == s)$stars
    degrees <- as.character(subset(data, scale == s)$degree)
    currCosts <- unique(subset(data, scale == s) %>% select(degree, cost))
    priors <- data %>% select(c(stars, prior.p)) %>% head(., n=5)

    results <- rbind(results,
                     data.frame(scale = s,
                                preds =
                                  as.numeric(mapply(computePosteriors[[calcFn]],
                                                    stars,
                                                    degrees,
                                                    MoreArgs =
                                                      list(m = mat,
                                                           alpha = alpha,
                                                           costs = currCosts,
                                                           priors = priors,
                                                           allDegrees = unique(degrees)))),
                                stars = stars,
                                degree = degrees,
                                stringsAsFactors = FALSE))
  }
  left_join(data, results)
}
```

Data set-up
```{r}
# All scales studied
# ------------------
scales <- c("good_excellent",
            "liked_loved",
            "palatable_delicious",
            "memorable_unforgettable",
            "some_all")
# Priors
# ------
unifPriors = data.frame(stars = seq(1, 5), prior.p = rep(0.2, 5))
empPriors = read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/emp_priors.csv")
yelpPriors <- data.frame(stars=1:5,
                         prior.p =
                           c(0.090190981, 0.105689431, 0.162783722, 0.304569543, 0.336766323))
# Cost / rewards
# --------------
emptyCosts5 <- data.frame(degree = c("low2", "low1", "mid", "hi2", "hi1"),
                            cost = rep(0.0, 5), stringsAsFactors = FALSE)
emptyCosts4 <- data.frame(degree = c("low2", "low1","hi2", "hi1"),
                            cost = rep(0.0, 4), stringsAsFactors = FALSE)
emptyCosts2 <- data.frame(degree = c("low", "hi"),
                            cost = rep(0.0, 2), stringsAsFactors = FALSE)
salienceCosts5 <- read.csv("~/Desktop/Projects/scalar_implicature/models/saliences5.csv")
salienceCosts4 <- read.csv("~/Desktop/Projects/scalar_implicature/models/saliences4.csv")
salienceCosts2 <- read.csv("~/Desktop/Projects/scalar_implicature/models/saliences2.csv")
normNegSalienceCosts <- function(saliences) {
  for (s in scales) {
    saliences[saliences$scale == s, ]$cost <-
      -1 * normedVec(saliences[saliences$scale == s, ]$cost)
  }
  saliences
}
# salience <- c(0.22, 0.08, 0.05, 1, 1)
# saliencePriors <- data.frame(stars=1:5, prior.p = (salience / sum(salience)))
# salienceCosts <- data.frame(degree = c("low2", "low1", "mid", "hi2", "hi1"),
#                             cost = -1 * (salience / sum(salience)), stringsAsFactors = FALSE)

# Speaker
# -------
speakerL12 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e12.csv")
speakerL10 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e10.csv")
speakerL08 <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L0_e8.csv")
# Listener
# --------
listener <- read.csv("~/Desktop/Projects/scalar_implicature/models/model_data/L1_e6.csv")

costs <- salienceCosts2
priors <- unifPriors
data <- left_join(speakerL08, costs) %>%
  left_join(., priors) %>%
  rowwise %>%
  select(scale, degree, stars, speaker.p, prior.p, cost) %>%
  group_by(scale)

# run fns()
checkSaliences()
tuneAlphas(modelData = data, humanDf = listener, numAlts = 2)
```

Data exploration Fns()
```{r}
# matchModelHumandData()
# -----------------------
# NEED TO FINISH WORKING HERE!!!!!!!!!!!!!!!!!!
matchModelHumanData <- function(predDf, listenerDf, study = "e12") {
  combinedDf <- data.frame(scale = c(),
                           degree = c(),
                           speaker.p = c(),
                           prior.p = c(),
                           cost = c(),
                           preds = c())

  for (s in unique(predDf$scale)) {
    if (study == "e12") {
      if (s == "some_all") {
        tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "mid")) %>%
          mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        } else {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        }
    } else if (study == "e10") {
        if (s == "some_all") {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "low1")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        } else {
          tempDf <- subset(predDf, scale == s & (degree == "hi1" | degree == "hi2")) %>%
            mutate(degree = ifelse(degree == "hi1", "hi", "low"))
        }
    } else { ## study e6
      tempDf <- subset(predDf, scale == s)
    }

    combinedDf <- rbind(combinedDf, tempDf)
  }

  results <- merge(x = combinedDf,
                   y = listenerDf,
                   by = c("degree", "stars", "scale"),
                   all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))

  results
}
# mergedData <- matchModelHumanData(data, listener, study = "e6")

# Alpha Tuning - maximixize correlation
# -------------------------------------
tuneAlphas <- function(modelData, humanDf, alphas = seq(1, 12), numAlts = 5, study = "e12"){
  sapply(alphas, function(i) {
    predDf <- runFull(modelData, calcFn = 2, alpha = i, numAlts = numAlts)
    mergedData <- matchModelHumanData(predDf, listener, study = study)
    cor(mergedData$preds, mergedData$listener.p)
  })
}
# tuneAlphas(modelData = data, humanDf = listener, numAlts = 4)

checkSaliences <- function(d, study = "e12", alpha = 1, numAlts = 5) {
  predDf <- runFull(d, calcFn = 2, alpha = alpha, numAlts = numAlts)
  mergedData <- matchModelHumanData(predDf, listener, study = study)
  cor(mergedData$preds, mergedData$listener.p)
}
# checkSaliences(data, study = "e6", alpha = 1, numAlts = 2)
```

Visualize
```{r}
mergedData <- matchModelHumanData(runFull(data, calcFn = 2, alpha = 7, numAlts = 5), listener)
plotData <- mergedData

# predPlot()
# ---------
# plot preds over stars with human data
predPlot <- function(d) {
  qplot(stars, listener.p, col=degree,
      data=d,
      main=paste("title")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = preds), lty = 4)
}
# correlationScatterPlot()
# -----------------------
# plot correleation with low / hi
correlationScatterPlot <- function(d) {
  ggplot(d, aes(x = listener.p, y = preds, col=stars)) +
  geom_point() + 
  stat_smooth(method="lm")
}
```

Entailment analysis
```{r}

```

Negative alts analysis
```{r}

```

Symmetric scale analysis
```{r}
# Set-up data

completeModelRun <- function(speaker, listener, priors, costs, numAlts) {
  results <- list() ## Will return this

  newCosts <- normNegSalienceCosts(costs)
  data <- left_join(speaker, newCosts) %>%
    left_join(., priors) %>%
    rowwise %>%
    select(scale, degree, stars, speaker.p, prior.p, cost) %>%
    group_by(scale)

  studyNum <- ifelse(numAlts == 5, "e12", ifelse(numAlts == 4, "e10", "e6"))
  # Tune alpha
  bestAlpha <- which.max(tuneAlphas(data, listener, numAlts = numAlts, study = studyNum))
  # Run model
  modelData <- runFull(data, calcFn = 2, alpha = bestAlpha, numAlts = numAlts)
  # Final data set for plotting / model fit

  combinedDf <- matchModelHumanData(modelData, listener, study = studyNum)
  fit <- cor(combinedDf$preds, combinedDf$listener.p)

  # plots
#   predPlot(combinedDf)
#   correlationScatterPlot(combinedDf)

  results[[1]] <- combinedDf
  results[[2]] <- modelData
  results[[3]] <- bestAlpha
  results[[4]] <- fit
  results[[5]] <- predPlot(combinedDf)
  results[[6]] <- correlationScatterPlot(combinedDf)
  results
}
```

Entailment / Neg alts / Symmetric scale
```{r}
symmetricScalerun <- completeModelRun(speakerL12, listener, unifPriors, salienceCosts5, numAlts = 5)
symmetricScalerun[[4]]
negAltsRun <-        completeModelRun(speakerL10, listener, unifPriors, salienceCosts4, numAlts = 4)
negAltsRun[[4]]
entailmentRun <-      completeModelRun(speakerL08, listener, unifPriors, salienceCosts2, numAlts = 2)
entailmentRun[[4]]
```

Aggregate analysis
```{r}
cors <- data.frame(cors = c(symmetricScalerun[[4]], negAltsRun[[4]], entailmentRun[[4]]),
                   m = c("full", "neg_alts", "entailment"))
ggplot(cors, aes(x=reorder(m, cors), y = cors)) +
  geom_bar(stat="identity", position="dodge") +
  ylab("Correlation") +
  xlab("Model") +
  ylim(0, 1) + 
  ggtitle("Model fit correlation coefficients") +
  guides(colour=FALSE) +
  geom_text(aes(x=m, y=as.numeric(cors) + 0.025, label=round(cors, 3)))
```

```{r}
normedVec(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
#normedVec(subset(test, scale == "good_excellent" & degree == "mid")$preds)
sum(subset(test, scale == "good_excellent" & degree == "hi1")$preds)
dfPreds <- test
dfHuman <- listener
name <- "good_excellent"
getCorrs <- function(dfPreds, dfHuman, name) {
  predSubset <- subset(dfPreds, scale == name & (degree == "hi1" | degree == "hi2"))
  predSubset$degree = with(predSubset,
                           ifelse(degree == "hi1", "hi", "low"))
  humanSubset <- subset(dfHuman, scale == name)
  joinedDf <- merge(x = predSubset,
                    y = humanSubset,
                    by = c("degree", "stars", "scale"),
                    all.x = TRUE) %>%
    mutate(listener.p = ifelse(is.na(listener.p), 0, listener.p))
  cor(joinedDf$preds, joinedDf$listener.p)
}

getCorrs(test, listener, "good_excellent")
getCorrs(test, listener, "liked_loved")
getCorrs(test, listener, "memorable_unforgettable")
getCorrs(test, listener, "palatable_delicious")

######################################################################
----------------------------------------------------------------------
######################################################################

runModel <- function(model, mat, degree, alpha) {
  if (model == 1) unlist(mapply(seq(1,5),
                                FUN=function(x) norm.post(x, degree, mat, alpha, priors)))
}
mat <- data %>% 
  select(stars, degree, speaker.p) %>%
  spread(degree, speaker.p) %>%
  mutate(hi1 = hi1 / sum(hi1),
         hi2 = hi2 / sum(hi2),
         mid = mid / sum(mid),
         low1 = low1 / sum(low1),
         low2 = low2 / sum(low2)) %>%
  select(hi1, hi2, mid, low1, low2)

runModel(1, mat[1:5, ], 1, 6)
runModel(1, mat[1:5, ], 2, 6)

subset(listener, scale == "good_excellent")

```

New cost formulation
```{r}
speaker.inf2 = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd2 = function(rating, degree, m, alpha, costs) {
  numerator = speaker.inf2(m[rating, degree], alpha, costs[degree])
  normalize = sum(unlist(sapply(seq(nrow(m)),
                                function(i) {speaker.inf2(m[i, ], alpha, costs[degree])})))
  return(numerator / normalize)
  # return(numerator)
}
# non-normalized posterior
# -----------------------
nn.post2 = function(rating, degree, m, alpha, costs, useprior) {
  prior = priors[rating, "prior.p"]
  # prior <- 0.2
  return(speaker.lhd2(rating, degree, m, alpha, costs) * prior)
}
# normalized posterior
# --------------------
norm.post2 = function(rating, degree, m, alpha, costs, useprior) {
  nn = nn.post2(rating, degree, m, alpha, costs, useprior)
  normalize = sum(unlist(sapply(seq(1, 5),
                                function(i){nn.post2(rating, i, m, alpha, costs, useprior)})))
  return(nn / normalize)
}
```

```{r}
# good excellent
mat[1:5,]
salience <- c(1, 1, 0.05, 0.08, 0.22)
normed.salience <- -1 * (salience / sum(salience))

normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[1:5,], 5, normed.salience))))
normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[1:5,], 5, normed.salience))))

#################################
# good_excellent
# --------------
# low
good.exc.low <-
  listener[with(listener, which(scale == "good_excellent" & degree == "low")),]$listener.p %>%
  c(. , rep(0, 2))
# high
good.exc.high <-
  listener[with(listener, which(scale == "good_excellent" & degree == "hi")),]$listener.p %>%
  c(. , rep(0, 3))
# new formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[1:5,], 7, normed.salience)))),
    good.exc.low[c(5, 4, 1, 2, 3)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[1:5,], 7, normed.salience)))),
    good.exc.high[c(5, 4, 3, 1, 2)])
# old formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 2, mat[1:5,], 7)))),
    good.exc.low[c(5, 4, 1, 2, 3)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 1, mat[1:5,], 7)))),
    good.exc.high[c(5, 4, 3, 1, 2)])

#################################
# liked_loved
# low
liked.loved.low <-
  listener[with(listener, which(scale == "liked_loved" & degree == "low")),]$listener.p %>%
  c(. , rep(0, 2))
# high
liked.loved.high <-
  listener[with(listener, which(scale == "liked_loved" & degree == "hi")),]$listener.p %>%
  c(. , rep(0, 3))
# new formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[6:10,], 7, normed.salience)))),
    liked.loved.low[c(5, 1, 2, 3, 4)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[6:10,], 7, normed.salience)))),
    liked.loved.high[c(5, 4, 3, 1, 2)])
# old formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 2, mat[6:10,], 7)))),
    liked.loved.low[c(5, 1, 2, 3, 4)])
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 1, mat[6:10,], 7)))),
    liked.loved.high[c(5, 4, 3, 1, 2)])

#################################
# memorable_unforgettable
# low
mem.unf.low <-
  listener[with(listener, which(scale == "memorable_unforgettable" & degree == "low")),]$listener.p %>%
  c(. , rep(0, 3))
mem.unf.low <- mem.unf.low[c(5, 4, 3, 2, 1)]
# high
mem.unf.high <-
  listener[with(listener,
                which(scale == "memorable_unforgettable" & degree == "hi")),]$listener.p %>%
  c(. , rep(0, 1))
mem.unf.high <- mem.unf.high[c(1, 5, 2, 3, 4)]
# new formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 2, mat[11:15,], 7, normed.salience)))),
    mem.unf.low)
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post2(x, 1, mat[6:10,], 7, normed.salience)))),
    mem.unf.high)
# old formulations
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 2, mat[6:10,], 3)))),
    mem.unf.low)
cor(normalize(unlist(mapply(seq(1,5),
                        FUN=function(x) norm.post(x, 1, mat[6:10,], 7)))),
    mem.unf.high)
```

How to implement this using saliency rewards (negative costs)???
```{r}
run.newCosts = function(d, alpha=1, useprior=F, usenone=F, normalize=F) {
  mat <- data.full.extras %>% 
    select(stars, degree, speaker.p) %>%
    spread(degree, speaker.p) %>%
    select(hi1, hi2, mid, low1, low2)

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  } 
  d$degree <- with(d, ifelse(degree == "hi1", 5,
                             ifelse(degree == "hi2", 4,
                                    ifelse(degree == "mid", 3, 
                                           ifelse(degree == "low1", 2, 3)))))
  d$pred = round(as.numeric(mapply(norm.post2, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha,
                                                    costs = normed.salience,
                                                    useprior = useprior))), 
                  digits=4)

  return(normalize(d))
}
run.newCosts(data.full.extras, alpha=7, useprior=TRUE, usenone=FALSE, normalize=normalize)
```

Old `run` function
```{r}
run.full = function(d, alpha=1, useprior=F, usenone=F, addMid=F, normalize=F) {
  # alpha = scales.entropy[scales.entropy$scale==d$scale[1], ]$Entropy
  if (addMid) {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               mid = mid / sum(mid),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, mid, low1, low2)    
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, mid, low1, low2)    
    }
  } else {
    if (normalize) {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        mutate(hi1 = hi1 / sum(hi1),
               hi2 = hi2 / sum(hi2),
               low1 = low1 / sum(low1),
               low2 = low2 / sum(low2)) %>%
        select(hi1, hi2, low1, low2)
    } else {
      mat = d %>%
        select(stars, degree, speaker.p) %>%
        spread(degree, speaker.p) %>%
        select(hi1, hi2, low1, low2)
    }
  }

  if (usenone) {
    mat$none = c(1, 0, 0, 0, 0)
  }

  d$pred = round(as.numeric(mapply(norm.post, d$stars, d$degree, 
                                    MoreArgs = list(m = mat, 
                                                    alpha = alpha, 
                                                    useprior = useprior))), 
                  digits=4)

  return(d)
}
```

```{r}
run.full(data.full.extras, useprior = TRUE, usenone = FALSE, addMid = TRUE, normalize = TRUE)
cor(data.full.extras$listener.p, data.full.extras$pred)
```