### 2016-01-26 analysis/BetweenExpDifferences.Rmd
initial between experiment tests for literal listener data

---
title: "BetweenExpDifferences"
author: "Ben Peloquin"
date: "January 25, 2016"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
    toc_depth: 2
---

```{r}
rm(list = ls())
setwd("/Users/benpeloquin/Desktop/Projects/scalar_implicature/analysis")
library(rjson)
library(jsonlite)
library(tidyr)
library(dplyr)
library(lme4)
library(lmerTest)
```

# Literal Listener Data

## Get all L0 listener data
```{r}

# readLiteralListenerData()
# filesPath :: path to directory with .json files ("../production-results/e8/")
readLiteralListenerData <- function(filesPath) {
  files <- dir(filesPath, pattern = "*.json")
  dRaw <- data.frame()

  for (f in files) {
    jf <- paste(filesPath, f, sep="")
    jd <- fromJSON(paste(readLines(jf), collapse=""))
    id <- data.frame(workerid = jd$WorkerId,
                     scale = jd$answers$data$scale,
                     degree = jd$answers$data$degree,
                     stars = jd$answers$data$manipulation_level,
                     judgment = as.numeric(jd$answers$data$judgment),
                     language = jd$answers$data$language)
    dRaw <- bind_rows(dRaw, id)
  }
  dRaw
}

e8 <- readLiteralListenerData("../production-results/e8/")
e10 <- readLiteralListenerData("../production-results/e10/")
e12 <- readLiteralListenerData("../production-results/e12/")
```

## Data cleaning and combine
```{r}
# workerid as factor
e8$workerid <- as.factor(e8$workerid)
e8$exp <- "e8"
e10$workerid <- as.factor(e10$workerid)
e10$exp <- "e10"
e12$workerid <- as.factor(e12$workerid)
e12$exp <- "e12"

# compliance checks here

```

## Predictor factoring
```{r}
allL0Listener <- rbind(e8, e10, e12)
allL0Listener$exp <- as.factor(allL0Listener$exp)
allL0Listener$judgment <- as.numeric(allL0Listener$judgment)
levels(allL0Listener$stars) <- c("5", "1", "2", "3", "4")
str(allL0Listener)
```

## glmer()
```{r}
# glmer WITHOUT experiment and with random effect for subjects
glm1 <- glmer(judgment ~ scale + degree + stars + (1 | workerid), data = allL0Listener)
# glmer WITH experiment and with random effect for subjects
glm2 <- glmer(judgment ~ exp + scale + degree + stars + (1 | workerid), data = allL0Listener)
```

## Visulaziations
```{r}
ggplot(allL0Listener, aes(x=stars, y=judgment, col=degree)) + 
  geom_point(alpha = 0.5) +
  geom_jitter(width = 0.1, height = 0.1) +
  facet_wrap(~scale + exp)
```

## Model Comaprisons
```{r}
summary(glm2)
anova(glm1, glm2)
```
Looks like we dont' see a significant improvement in model fit when we include the fixed effect for experiment. However, it does look like there might be a significant difference

Inidivual words
```{r}
if (scale == "some_all") {
  if (exp == "10") {
    if (degree == "hi1") {
      d$word <- "most"
    } else if (degree == "low1") {
      d$word <- "some"
    } else 
  }
}

if (scale == "some_all") {
  switch(degree, "hi2" = "all", "hi1" = "most", "low1" = "some")
}
a <- data.frame(scale = rep("some_all", 4), degree = c("hi2", "hi1", "low1", "low2"), stringsAsFactors = FALSE)

for (d in a$degree) {
  print(d)
  a$word <- switch(d, "hi2" = "all", "hi1" = "most", "low1" = "some", "low2" = "none")  
}

a
```