### 2016-01-26 analysis/BetweenExpDifferences.Rmd
included specific scalr words, improved compliance checks

---
title: "BetweenExpDifferences"
author: "Ben Peloquin"
date: "January 25, 2016"
output:
  html_document:
    highlight: pygments
    theme: flatly
    toc: yes
    toc_depth: 2
---

```{r}
rm(list = ls())
setwd("/Users/benpeloquin/Desktop/Projects/scalar_implicature/analysis")
library(rjson)
library(jsonlite)
library(tidyr)
library(dplyr)
library(lme4)
library(lmerTest)
```

# Literal Listener Data

## Get all L0 listener data
```{r}

# readLiteralListenerData()
# filesPath :: path to directory with .json files ("../production-results/e8/")
readLiteralListenerData <- function(filesPath) {
  files <- dir(filesPath, pattern = "*.json")
  dRaw <- data.frame()

  for (f in files) {
    jf <- paste(filesPath, f, sep="")
    jd <- fromJSON(paste(readLines(jf), collapse=""))
    id <- data.frame(workerid = jd$WorkerId,
                     scale = jd$answers$data$scale,
                     degree = jd$answers$data$degree,
                     stars = jd$answers$data$manipulation_level,
                     judgment = as.numeric(jd$answers$data$judgment),
                     language = jd$answers$data$language)
    dRaw <- bind_rows(dRaw, id)
  }
  dRaw
}

e8 <- readLiteralListenerData("../production-results/e8/")
e10 <- readLiteralListenerData("../production-results/e10/")
e12 <- readLiteralListenerData("../production-results/e12/")
```

## Data org / cleaning
```{r}
# workerid as factor
e8$workerid <- as.factor(e8$workerid)
e8$exp <- "e8"
e10$workerid <- as.factor(e10$workerid)
e10$exp <- "e10"
e12$workerid <- as.factor(e12$workerid)
e12$exp <- "e12"

# compliance checks here
getCompliantWorkerIds <- function(df, exp) {
  if (exp == "e8") {
    compliant <- df %>% 
      filter(tolower(language) == "english") %>%
      filter(grepl("training1", scale)) %>%
      group_by(workerid) %>%
      summarise(compliant = judgment[scale == "training1" & degree == "hi"] == 1 & 
                  judgment[scale == "training1" & degree == "low"] == 0)
  } else {
    compliant <- df %>% 
      filter(tolower(language) == "english") %>%
      filter(grepl("training", scale)) %>%
      group_by(workerid) %>%
      summarise(compliant = (judgment[scale == "training1" & degree == "hi2"] == 1) && 
              (judgment[scale == "training1" & degree == "low1"] == 1))
  }
  # return compliant ids
  compliant
}

# Append compliance booleans
e8CompliantIds <- getCompliantWorkerIds(e8, "e8")
e8 <- left_join(e8, e8CompliantIds)
e10CompliantIds <- getCompliantWorkerIds(e10, "e10")
e10 <- left_join(e10, e10CompliantIds)
e12CompliantIds <- getCompliantWorkerIds(e12, "e12")
e12 <- left_join(e12, e12CompliantIds)
```

# Populate literal listener combined data
```{r}
# filter out compliance and training
allL0Listener <- rbind(e8, e10, e12) %>%
  subset(scale != "training1" & scale != "training2") %>%
  filter(compliant == TRUE)

allL0Listener$exp <- as.factor(allL0Listener$exp) ## exp. is factor
allL0Listener$judgment <- as.numeric(allL0Listener$judgment) ## judgment is numeric
levels(allL0Listener$stars) <- c("5", "1", "2", "3", "4") ## re-arrange stars
str(allL0Listener)
```

## Add inidivual words
```{r}
lookupScalar <- function(degree, scale, exp) {
  # some_all
  if (scale == "some_all") {
    if (exp == "e12") {
      if (degree == "hi1") return("all")
      if (degree == "hi2") return("most")
      if (degree == "mid") return("some")
      if (degree == "low1") return("little")
      if (degree == "low2") return("none")
    } else if (exp == "e10") {
      if (degree == "hi1") return("all")
      if (degree == "hi2") return("most")
      if (degree == "low1") return("some")
      if (degree == "low2") return("none")
    } else { ## must be exp 8
      if (degree == "hi") return("all")
      if (degree == "low") return("some")
    }
  # liked_loved
  } else if (scale == "liked_loved") {
    if (degree == "hi1" | degree == "hi") return("loved")
    if (degree == "hi2" | degree == "low") return("liked")
    if (degree == "mid") return("indifferent")
    if (degree == "low1") return("disliked")
    if (degree == "low2") return("hated")
  # good_excellent
  } else if (scale == "good_excellent") {
    if (degree == "hi1" | degree == "hi") return("excellent")
    if (degree == "hi2" | degree == "low") return("good")
    if (degree == "mid") return("okay")
    if (degree == "low1") return("bad")
    if (degree == "low2") return("horrible")
  # paltable_delicious
  } else if (scale == "palatable_delicious") {
    if (degree == "hi1" | degree == "hi") return("delicious")
    if (degree == "hi2" | degree == "low") return("palatable")
    if (degree == "mid") return("mediocre")
    if (degree == "low1") return("gross")
    if (degree == "low2") return("disgusting")
  # memorable_unforgettable
  } else if (scale == "memorable_unforgettable") {
    if (degree == "hi1" | degree == "hi") return("unforgettable")
    if (degree == "hi2" | degree == "low") return("memorable")
    if (degree == "mid") return("ordinary")
    if (degree == "low1") return("bland")
    if (degree == "low2") return("forgettable")
  }
  NA
}

allL0Listener$word <- as.factor(with(allL0Listener, mapply(lookupScalar, degree, scale, exp)))
```

## glmer()
```{r}
# fix contrasts so we're comparing most disparate experimetns (e8 and e12)
contrasts(allL0Listener$exp) <- cbind("e10" = c(1, 0, 0), "e8" = c(0, 0, 1))

# glmer WITHOUT experiment and with random effect for subjects
glm1 <- glmer(judgment ~ scale + word + stars + (1 | workerid), data = allL0Listener)
# glmer WITH experiment and with random effect for subjects
glm2 <- glmer(judgment ~ exp + word + scale +  stars + (1 | workerid), data = allL0Listener)
```

## Visulaziations?
```{r}
# ggplot(allL0Listener, aes(x=stars, y=judgment, col=degree)) + 
#   geom_point(alpha = 0.5) +
#   geom_jitter(width = 0.1, height = 0.1) +
#   facet_wrap(~scale + exp)
```

## Model Comaprisons
```{r}
summary(glm2)
anova(glm1, glm2)
```
Looks like we dont' see a significant improvement in model fit when we include the fixed effect for experiment. However, it does look like there might be a significant difference